### 2010-04-01 main/doc/acknowledgments.tex

\section{Acknowledgments}
\label{sec:acknowledgments}

Chord would not be possible without the following open-source software:

\begin{itemize}
\item
\href{http://joeq.sourceforge.net/}{Joeq}, a Java compiler framework
\item
\href{http://www.csg.is.titech.ac.jp/~chiba/javassist/}{Javassist},
a Java bytecode manipulation tool
\item
\href{ttp://bddbddb.sourceforge.net/}{bddbddb}, a BDD-based Datalog solver
\end{itemize}

Chord additionally relies on the following open-source tools and libraries:
\begin{itemize}
\item
\href{http://ant-contrib.sourceforge.net/}{Ant-Contrib},
a collection of useful Ant tasks  
\item
\href{http://buddy.sourceforge.net/}{BuDDy}, a BDD library
\item
\href{http://trove4j.sourceforge.net/}{GNU Trove},
a primitive collections library for Java
\item
\href{http://www.java2html.com/}{Java2HTML} and
\href{http://www.java2html.de/}{Java2Html}, Java to HTML tools
\item
\href{http://saxon.sourceforge.net/}{Saxon}, an XSLT processor
\end{itemize}

Chord was supported in part by grants from the National Science Foundation,
an equipment grant from Intel, and a Microsoft fellowship during 2005-2007.