### 2010-04-01 main/doc/deadlock.tex

\section{Deadlock Analysis}
\label{sec:deadlock}

Run the deadlock analysis provided in Chord by running the following command in Chord's \verb+main/+ directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
\end{verbatim}
\end{quote}

where directory \verb+<...>+ contains a file named \verb+chord.properties+ which 
defines properties \verb+chord.main.class+, \verb+chord.class.path+, and \verb+chord.src.path+.
See Section \ref{sec:chord-sysprops} for the meaning of these properties.

Directory \verb+main/examples/deadlock_test/+ provides a toy Java program on which you can run the deadlock
analysis by first running \verb+ant+ in that directory (in order to compile the program's \verb+.java+ files to
\verb+.class+ files) and then running the following command in Chord's \verb+main/+ directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=examples/deadlock_test/ -Dchord.run.analyses=deadlock-java run
\end{verbatim}
\end{quote}

Upon successful completion, the file \verb+deadlocks.html+ should be produced in directory
\verb+main/examples/deadlock_test/chord_output/+.