### 2010-04-01 main/doc/running_analysis.tex

\section{Running a Program Analysis}
\label{sec:running-analysis}

TODO

%chord.run.analyses
%List of names of program analyses to be run in order.
%chord.java.analysis.path	 path	 Classpath containing program analyses in Java (i.e. @Chord-annotated classes).	${chord.main.dir}/classes/
%chord.dlog.analysis.path	 path	 Path of directories containing program analyses in Datalog (i.e. *.datalog and *.dlog files).	${chord.main.dir}/src/main/dlog/
%chord.reuse.rels	 bool	 Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.	false
%chord.publish.results	 bool	 Publish the results of program analyses in HTML.	true