### 2010-04-01 main/doc/writing_analysis.tex

\section{Writing a Program Analysis}
\label{sec:writing-analysis}

TODO