### 2010-04-02 main/doc/chord_sysprops.tex

\section{Chord System Properties}
\label{sec:chord-sysprops}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.

\begin{enumerate}
\item
\code{chord.main.dir} \\
{\bf Type:} location \\
{\bf Description:} \code{main/} directory in Chord's installation.

\item
\code{chord.lib.dir} \\
{\bf Type:} location \\
{\bf Description:} Directory containing libraries needed by Chord. \\
{\bf Default value:} \code{{chord.main.dir}/lib/}

\item
\code{chord.main.class.path} \\
{\bf Type:} path \\
{\bf Description:} Classpath of Chord.  It includes the path specified by property \code{chord.java.analysis.path} to allow running user-defined program analyses. \\
{\bf Default value:} See \code{{chord.main.dir}/build.xml}

\item
\code{chord.bddbddb.class.path} \\
{\bf Type:} path \\
{\bf Description:} Classpath of \code{bddbddb}. \\
{\bf Default value:} See \code{{chord.main.dir}/build.xml}

\item
\code{chord.instr.agent.file} \\
{\bf Type:} location \\
{\bf Description:} Shared library implementing Chord instrumentation agent. \\
{\bf Default value:} \code{{chord.main.dir}/lib/libchord_instr_agent.so}

\item
\code{chord.javadoc.url} \\
{\bf Type:} string \\
{\bf Description:} URL of the Javadoc location of program analyses.  It is used when publishing targets (i.e., when property \code{chord.publish.targets} is set to true). \\
{\bf Default value:} \code{http://chord.stanford.edu/javadoc/}

\item
\code{chord.work.dir} \\
{\bf Type:} location \\
{\bf Description:} Working directory during Chord's execution. \\
{\bf Default value:} current working directory

\item
\code{chord.props.file} \\
{\bf Type:} location \\
{\bf Description:} Properties file loaded by \code{{chord.main.dir}/build.xml}.  Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g., for user-defined analyses), may be set in this file.  Each relative (instead of absolute) path element in the value of any property named \code{chord.<...>.path} set in this file is converted to an absolute path element with respect to the directory containing this file. \\
{\bf Default value:} \code{{chord.work.dir}/chord.properties}

\item
\code{chord.max.heap} \\
{\bf Type:} string \\
{\bf Description:} Maximum memory size of JVM running Chord. \\
{\bf Default value:} \code{1024m}

\item 
\code{chord.max.stack} \\
{\bf Type:} string \\
{\bf Description:} Maximum thread stack size of JVM running Chord. \\
{\bf Default value:} \code{32m}

\item
\code{chord.jvmargs} \\
{\bf Type:} string \\
{\bf Description:} Arguments passed to JVM running Chord. \\
{\bf Default value:} \code{"-showversion -ea -Xmx{chord.max.heap} -Xss{chord.max.stack}"}

\item
\code{chord.bddbddb.max.heap} \\
{\bf Type:} string \\
{\bf Description:} Maximum memory size of JVM running \bddbddb. \\
{\bf Default value:} \code{1024m}

\item
\code{chord.build.scope} \\
 bool \\
 Compute analysis scope (i.e., reachable classes and methods)
using algorithm specified by properties \code{chord.scope.kind} and \code{chord.reuse.scope}.
Program scope is computed regardless of the value of this property if another task (e.g., a
program analysis to be run) needs it.
{\bf Default value:} \code{false} \\

\item
\code{chord.run.analyses} \\
 string list \\
 List of names of program analyses to be run
in order.
{\bf Default value:} \code{""}

\item
\code{chord.print.rels} \\
{\bf Type:} string list \\
{\bf Description:} List of names of program relations whose contents must be printed to files \code{{chord.out.dir}/<...>.txt} where \code{<...>} denotes the relation name.  \\
{\bf Default value:} \code{""}

\item
\code{chord.publish.targets} \\
 bool \\
 Publish all targets defined by program analyses in
paths specified by properties \code{chord.java.analysis.path} and
\code{chord.dlog.analysis.path}.
{\bf Default value:} \code{false} \\

\item
\code{chord.main.class} \\
 class \\
 Fully-qualified name of main class of given program
(e.g., \code{com.example.Main}). \\

\item
\code{chord.class.path} \\
{\bf Type:} path \\
 Classpath of given program. \code{{sun.boot.class.path}}
is implicitly appended to it.
{\bf Default value:} \code{{sun.boot.class.path}}

\item
\code{chord.src.path} \\
{\bf Type:} path \\
 Java source path of given program. \\
{\bf Default value:} \code{""}

\item
\code{chord.run.ids} \\
 string list \\
 List of IDs to identify runs of given program.
{\bf Default value:} \code{0}

\item
\code{chord.args.XXX} \\
 string \\
 Command-line argument string to be used for given
program in run having ID \code{XXX}.
{\bf Default value:} \code{""}

\item
\code{chord.runtime.jvmargs} \\
{\bf Type:} string \\
{\bf Description:} Arguments to JVM when running given program.  It is used when the program is executed to compute its scope (i.e., when property \code{chord.scope.kind} is set to \code{dynamic}) as well as when the instrumented program is executed for dynamic analyses. \\
{\bf Default value:} \code{"-ea -Xmx1024m"}

\item
\code{chord.scope.kind} \\
{\bf Type:} [\code{rta} $|$ \code{dynamic}] \\
{\bf Description:} Algorithm to compute analysis scope (i.e., reachable classes and methods).  Current choices include \code{rta} (static Rapid Type Analysis) and \code{dynamic} (dynamic analysis). \\
{\bf Default value:} \code{rta}

\item
\code{chord.reuse.scope} \\
 bool \\
 Treat analysis scope as the classes and methods listed
in files specified by properties \code{chord.classes.file} and \code{chord.methods.file},
respectively. Property \code{chord.scope.kind} is ignored if this property is set to
\code{true} and the two files exist.
{\bf Default value:} \code{false}

\item
\code{chord.check.exclude} \\
 string list \\
 List of prefixes of names of classes and
packages to be excluded by program analyses.
Interpretation of this property is analysis-specific. \\
{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."} \\

\item
\code{chord.instr.exclude} \\
{\bf Type:} string list \\
{\bf Description:} List of prefixes of names of classes and packages to be excluded from being instrumented.  Classes in package \code{java.lang.ref} are not instrumented regardless of the value of this property. \\
{\bf Default value:} \code{"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."}

\item
\code{chord.java.analysis.path} \\
{\bf Type:} path \\
{\bf Description:} Classpath containing program analyses written in Java (i.e., \code{@Chord}-annotated classes). \\
{\bf Default value:} \code{{chord.main.dir}/classes/}

\item
\code{chord.dlog.analysis.path} \\
{\bf Type:} path \\
{\bf Description:} Path of directories containing program analyses written in Datalog (i.e., \code{*.datalog} and \code{*.dlog} files). \\
{\bf Default value:} \code{{chord.main.dir}/src/dlog/}

\item
\code{chord.reuse.rels} \\
 bool \\
 Construct program relations from BDDs stored on disk
(from a previous run of Chord) whenever possible instead of re-computing them.
{\bf Default value:} \code{false} \\
\item
\code{chord.publish.results} \\
 bool \\
 Publish the results of program analyses in HTML.
Interpretation of this property is analysis-specific.
{\bf Default value:} \code{true} \\

\item
\code{chord.ssa} \\
 bool \\
 Do SSA transformation for all methods deemed reachable by the
algorithm used to compute analysis scope.
{\bf Default value:} \code{true} \\

\item
\code{chord.bddbddb.noisy} \\
 bool \\
 Make \bddbddb\ output debug info when solving Datalog programs.
{\bf Default value:} \code{false} \\

\item
\code{chord.save.maps} \\
 bool \\
 Write to file \code{{chord.bddbddb.work.dir}/<...>.map}
when saving program domain named \code{<...>} (useful for debugging Datalog programs using
\code{debug} target of \code{{chord.main.dir}/build.xml}).
{\bf Default value:} \code{true} \\

\item
\code{chord.trace.pipe} \\
 bool \\
 Implement the dynamic trace file as a POSIX pipe instead of a regular file.
{\bf Default value:} \code{true} \\

\item
\code{chord.trace.block.size} \\
{\bf Type:} int \\
{\bf Description:} Number of bytes to read/write in a single operation from/to the dynamic trace file.
{\bf Default value:} \code{4096} \\

\item
\code{chord.runtime.class} \\
{\bf Type:} class \\
{\bf Description:} Subclass of \code{chord.project.Runtime} used by dynamic program analyses to handle events generated during an instrumented program's execution. \\
{\bf Default value:} \code{chord.project.BufferedRuntime}

\item
\code{chord.out.dir} \\
{\bf Type:} location \\
{\bf Description:} Directory to which Chord dumps all files. \\
{\bf Default value:} \code{{chord.work.dir}/chord_output/}

\item
\code{chord.out.file} \\
{\bf Type:} location \\
{\bf Description:} File to redirect standard output stream during Chord's execution. \\
{\bf Default value:} \code{{chord.out.dir}/log.txt}

\item
\code{chord.err.file} \\
{\bf Type:} location \\
{\bf Description:} File to redirect standard error stream during Chord's execution. \\
{\bf Default value:} \code{{chord.out.dir}/log.txt}

\item
\code{chord.classes.file} \\
{\bf Type:} location \\
{\bf Description:} File from/to which list of classes deemed reachable is read/written. \\
{\bf Default value:} \code{{chord.out.dir}/classes.txt}

\item
\code{chord.methods.file} \\
{\bf Type:} location \\
{\bf Description:} File from/to which list of methods deemed reachable is read/written. \\
{\bf Default value:} \code{{chord.out.dir}/methods.txt}

\item
\code{chord.bddbddb.work.dir} \\
{\bf Type:} location \\
{\bf Description:} Directory used by \bddbddb\ as its input/output directory. \\
{\bf Default value:} \code{{chord.out.dir}/bddbddb/}

\item
\code{chord.boot.classes.dir} \\
{\bf Type:} location \\
{\bf Description:} Directory containing instrumented JDK classes used by given program. \\
{\bf Default value:} \code{{chord.out.dir}/boot_classes/}

\item
\code{chord.user.classes.dir} \\
{\bf Type:} location \\
{\bf Description:} Directory containing instrumented non-JDK classes of given program. \\
{\bf Default value:} \code{{chord.out.dir}/user_classes/}

\item
\code{chord.instr.scheme.file} \\
{\bf Type:} location \\
{\bf Description:} File specifying kind and format of events in dynamic trace file. \\
{\bf Default value:} \code{{chord.out.dir}/scheme.ser}

\item
\code{chord.crude.trace.file} \\
{\bf Type:} location \\
{\bf Description:} Crude dynamic trace file. \\
{\bf Default value:} \code{{chord.out.dir}/crude_trace.txt}

\item
\code{chord.final.trace.file} \\
{\bf Type:} location \\
{\bf Description:} Final dynamic trace file. \\
{\bf Default value:} \code{{chord.out.dir}/final_trace.txt}
\end{enumerate}