### 2010-04-02 main/doc/writing_dynamic_analysis.tex

\section{Writing a Dynamic Program Analysis}
\label{sec:writing-dynamic-analysis}

Follow the following steps to write your own dynamic analysis.

{\bf Step 1: Setting up the Instrumentation Scheme}

Determine the instrumentation scheme required by your dynamic analysis, that is,
the kind and format of events to be generated during an instrumented program's
execution. See Section \ref{sec:instr-events} 
for the kinds of supported events and their formats.

{\bf Step 2: Implementing the Analysis}

Create a class extending \code{chord.project.DynamicAnalysis} and override
the appropriate methods in it.
The only method that must be compulsorily overridden is method \code{getInstrScheme()},
which must return an instance of the instrumentation scheme chosen in Step 1 above,
plus each \code{process<event>(<args>)} method that corresponds to event \code{<event>}
with format \code{<args>} enabled by the chosen instrumentation scheme.
All other methods are no-ops if not overridden.

A sample such class called \code{MyDynamicAnalysis} is shown below:

\begin{verbatim}
    import chord.project.DynamicAnalysis;
    import chord.instr.InstrScheme;

    // ***TODO***: analysis won't be recognized by Chord without this annotation
    @Chord(name=<name-of-analysis>)

    public class MyDynamicAnalysis extends DynamicAnalysis {
        InstrScheme scheme;

        public InstrScheme getInstrScheme() {
            if (scheme != null)
                return scheme;
            scheme = new InstrScheme();
            // ***TODO***: Choose (<event1>, <args1>), ... (<eventN>, <argsN>)
            // depending upon the kind and format of events required by this
            // dynamic analysis to be generated for this during an instrumented
            // program's execution.
            scheme.set<event1>(<args1>);
            ...
            scheme.set<eventN>(<argsN>);
            return scheme;
        }

        public void initAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // before all instrumented program runs start.
        }

        public void doneAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // after all instrumented program runs finish.
        }

        public void initPass() {
            // ***TODO***: User code to be executed once before each
            // instrumented program run starts.
        }

        public void donePass() {
            // ***TODO***: User code to be executed once after each
            // instrumented program run finishes.
        }

        // User-defined event handlers for this dynamic analysis.
        // No-ops if not overridden.
        public void process<event1>(<args1>) {
            // ***TODO***: User code for handling events of kind
            // <event1> with format <args1>.
        }
        ...
        public void process<eventN>(<argsN>) {
            // ***TODO***: User code for handling events of kind
            // <eventN> with format <argsN>.
        }
    }
\end{verbatim}

{\bf Step 3: Choosing or Implementing the Runtime Event Handler}

Determine the runtime event handler for your dynamic analysis. There are two kinds of runtime event
handlers, offline and online; see package \code{chord.runtime} for more details.

For most dynamic analyses, an offline event hander suffices, in which case you don't need to do
anything in this step.
Otherwise, you need to create a class extending \code{chord.runtime.Runtime} and override the
appropriate methods in it, and provide the fully-qualified name of the created class as the value of property
\code{chord.runtime.class} (whose default value is \code{chord.runtime.BufferedRuntime}).

{\bf Step 4: Compiling the Analysis}

Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
above in the path defined by property \code{chord.java.analysis.path}.

{\bf Step 5: Specifying Program Inputs}

Provide the IDs of program runs to be generated (say 1, 2, ..., M) and the command-line arguments to be
used for the program in each of those runs (say \code{<args1>}, ..., \code{<argsM>}) via properties
\code{chord.run.ids=1,2,...,N} and \code{chord.args.1=<args1>}, ..., \code{chord.args.M=<argsM>}.
By default, \code{chord.run.ids=0} and \code{chord.args.0=""}, that is, the program will be run only
once (using run ID 0) with no command-line arguments.

{\bf Step 6: Running the Analysis}

To run the analysis, set property \code{chord.run.analyses} to \code{<name-of-analysis>}
(recall that \code{<name-of-analysis>} is the name provided in the \code{@Chord} annotation for class
\code{MyDynamicAnalysis} created above).

%(except that you may want to change the default values of properties
%\code{chord.trace.pipe} and \code{chord.trace.block.size} for performance, or the default values
%of properties \code{chord.trace.pipe} and \code{chord.reuse.trace} for determinism across; see
%Section \ref{sec:chord-sysprops} for the meaning of these properties).