### 2010-04-02 main/doc/instr_events.tex

\section{Instrumentation Events}
\label{sec:instr-events}

Chord allows a given Java program to be instrumented to generate the
following kinds of events during its execution.

\begin{enumerate}
\item
Method entry and exit. \\
{\bf Event:} \code{ENTER_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated after thread $t$ enters method $m$ respectively. \\
{\bf Event:} \code{LEAVE_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated before thread $t$ leaves method $m$ respectively.

\item
Loop entry and exit. \\
{\bf Event:} \code{ENTER_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated before thread $t$ enters loop $w$ respectively. \\
{\bf Event:} \code{LEAVE_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated after thread $t$ leaves loop $w$ respectively.

\item
Object allocation \\
{\bf Event:} \code{BEF_NEW} $h$ $t$ \\
{\bf Desc.:} Generated only in the crude trace before thread $t$ executes a \code{new} bytecode instruction at program point $h$. \\
{\bf Event:} \code{AFT_NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated only in the crude trace after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
%Note: Both above events are visible only in the crude trace and not in the final trace. They are generated only if argument \code{hasObj} of method {@link #setNewAndNewArrayEvent} is true.
{\bf Event:} \code{NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
{\bf Event:} \code{NEW_ARRAY} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{newarray} bytecode instruction at program point $h$ and allocates new object $o$.

\item
Getstatic primitive \\
{\bf Event:} \code{GETSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed static field $f$ at program point $e$.

\item
Getstatic reference \\
{\bf Event:} \code{GETSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed static field $f$ at program point $e$.

\item
Putstatic primitive \\
{\bf Event:} \code{PUTSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed static field $f$ at program point $e$.

\item
Putstatic reference \\
{\bf Event:} \code{PUTSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed static field $f$ at program point $e$.

\item
Getfield primitive \\
{\bf Event:} \code{GETFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Getfield reference \\
{\bf Event:} \code{GETFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield primitive \\
{\bf Event:} \code{PUTFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield reference \\
{\bf Event:} \code{PUTFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Aload primitive \\
{\bf Event:} \code{ALOAD_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ reads the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Aload reference \\
{\bf Event:} \code{ALOAD_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore primitive \\
{\bf Event:} \code{ASTORE_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ writes the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore reference \\
{\bf Event:} \code{ASTORE_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Method call \\
{\bf Event:} \code{METHOD_CALL} $i$ $t$ \\
{\bf Desc.:} Generated before or after thread $t$ executes the method invocation statement at program point $i$.

\item
Thread start call \\
{\bf Event:} \code{THREAD_START} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{start()} method of \code{java.lang.Thread} at program point $i$ and spawns a thread $o$.

\item
Thread join call \\
{\bf Event:} \code{THREAD_JOIN} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread $o$.

\item
Lock acquire \\
{\bf Event:} \code{ACQUIRE_LOCK} $l$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a statement of the form monitorenter $o$ or enters a method synchronized on $o$ at program point $l$.

\item
Lock release \\
{\bf Event:} \code{RELEASE_LOCK} $r$ $t$ $o$ \\
{\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.

\item
Thread wait call \\
{\bf Event:} \code{WAIT} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{wait()} method of \code{java.lang.Object} at program point $i$ on object $o$.

\item
Thread notify call \\
{\bf Event:} \code{NOTIFY} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point $i$ on object $o$.
\end{enumerate}