### 2010-08-15 main/doc/deadlock.tex
minor (added documentation + fixed javadoc tags)

\section{Deadlock Analysis}
\label{sec:deadlock}

Run the deadlock analysis provided in Chord by running the following command in Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
\end{verbatim}

\noindent where directory {\tt <...>} contains a file named \code{chord.properties} which
defines properties \code{chord.main.class}, \code{chord.class.path}, and \code{chord.src.path}.
See Section \ref{sec:chord-sysprops} for the meaning of these properties.

Directory \code{main/examples/deadlock_test/} provides a toy Java program on which you can run
the deadlock analysis.
First run {\tt ant} in that directory (in order to compile the program's {\tt .java} files to
{\tt .class} files) and then run the above command in Chord's {\tt main/} directory with
{\tt <...>} replaced by \code{examples/deadlock_test/}.
Upon successful completion, the file \code{deadlocks.html} should be produced in directory
\code{main/examples/deadlock_test/chord_output/}.