### 2010-12-13 main/doc/program_setup.tex

\section{Setting up a Java Program}
\label{sec:program-setup}

Suppose the program to be analyzed has the following directory structure:

\begin{verbatim}
    example/
        src/
            foo/
                Main.java
                ...
        classes/
            foo/
                Main.class
                ...
        lib/
            src/
                taz/
                    ...
            jar/
                taz.jar
        chord.properties
\end{verbatim}

\noindent The above structure is typical: the program's Java source
files are under {\tt src/}, its class files are under {\tt classes/},
the source and jar files of the libraries used by the program are
under \code{lib/src/} and \code{lib/jar/}, respectively.

File \code{chord.properties} specifies properties to be passed to
Chord (alternatively, these properties may be passed on the
command-line, in format \code{-D<name>=<value>}).  A sample such file
for the above program is as follows:

\begin{verbatim}
    chord.main.class=foo.Main
    chord.class.path=classes:lib/jar/taz.jar
    chord.src.path=src:../lib/src
    chord.run.ids=0,1
    chord.args.0="-thread 1 -n 10"
    chord.args.1="-thread 2 -n 50"
\end{verbatim}

Each relative path element in the value of any property named
\code{chord.<...>.path} defined in this file is converted to an
absolute path element with respect to the directory containing this
file, which in the above case is \code{example/}.

Section \ref{sec:chord-sysprops} presents an exhaustive listing of
properties recognized by \Chord.  Here, we only describe those defined
in the above sample \code{chord.properties} file:

\begin{itemize}
\item
\code{chord.main.class} specifies the fully-qualified name of the main
class of the program.
\item
\code{chord.class.path} specifies the application-specific classpath
of the program (the standard library classpath is implicitly
included).
\item
\code{chord.src.path} specifies the Java source path of the program.
All program analyses in Chord operate on Java bytecode.  The only use
of this property is to HTMLize the Java source files of the program so
that the results of program analyses can be reported at the Java
source code level.
\item
\code{chord.run.ids} specifies a list of IDs to identify runs of the
program.  It is used by dynamic program analyses to determine how many
times the program must be run.  An additional use of this property is
to allow specifying the command-line arguments to use in the run
having ID {\tt <id>} via property \code{chord.args.<id>}, as
illustrated by properties \code{chord.args.0} and \code{chord.args.1}
in the above example.
\end{itemize}

To run Chord on the above program, run the following command in
Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...>/example run
\end{verbatim}

\noindent where {\tt <...>} denotes the absolute path of the parent of
directory {\tt example/}.  Property \code{chord.work.dir} specifies
the working directory during Chord's execution.

The above command causes Chord to load the \code{chord.properties}
file if present.  The location of this file may be specified
explicitly via property \code{chord.props.file} on the above command
line.  By default, it is \code{[chord.work.dir]/chord.properties}.

In the above case, Chord does not do much beyond loading the sample
\code{chord.properties} file.  For Chord to do something interesting,
you need to set additional properties, either on the above command
line or in the \code{chord.properties} file, specifying the task(s)
Chord must perform.  All tasks are summarized in Section
\ref{sec:chord-sysprops:task}.  The two most common tasks, described
next, are computing the scope of the given program (Section
\ref{sec:building-scope}) and running analyses on the given program
(Section \ref{sec:running-analysis}).