### 2010-12-18 main/doc/instr_events.tex

\section{Instrumentation events}
\label{sec:instr-events}

Chord allows a given Java program to be instrumented to generate the
following kinds of events during its execution.

\begin{enumerate}
\item
Method entry and exit. \\
{\bf Event:} \code{ENTER_METHOD} m t \\
{\bf Desc.:} Generated after thread t enters method m respectively. \\
{\bf Event:} \code{LEAVE_METHOD} m t \\
{\bf Desc.:} Generated before thread t leaves method m respectively.

\item
Loop entry and exit. \\
{\bf Event:} \code{ENTER_LOOP} w t \\
{\bf Desc.:} Generated before thread t enters loop w respectively. \\
{\bf Event:} \code{LEAVE_LOOP} w t \\
{\bf Desc.:} Generated after thread t leaves loop w respectively.

\item
Object allocation \\
{\bf Event:} \code{BEF_NEW} h t \\
{\bf Desc.:} Generated before thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
{\bf Event:} \code{AFT_NEW} h t o \\
{\bf Desc.:} Generated after thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
{\bf Event:} \code{NEW_ARRAY} h t o \\
{\bf Desc.:} Generated after thread t executes a \code{newarray} bytecode instruction at program point h and allocates new object o.

\item
Getstatic primitive \\
{\bf Event:} \code{GETSTATIC_PRIMITIVE} e t f \\
{\bf Desc.:} Generated after thread t reads primitive-typed static field f at program point e.

\item
Getstatic reference \\
{\bf Event:} \code{GETSTATIC_REFERENCE} e t f o \\
{\bf Desc.:} Generated after thread t reads object o from reference-typed static field f at program point e.

\item
Putstatic primitive \\
{\bf Event:} \code{PUTSTATIC_PRIMITIVE} e t f \\
{\bf Desc.:} Generated after thread t writes primitive-typed static field f at program point e.

\item
Putstatic reference \\
{\bf Event:} \code{PUTSTATIC_REFERENCE} e t f o \\
{\bf Desc.:} Generated after thread t writes object o to reference-typed static field f at program point e.

\item
Getfield primitive \\
{\bf Event:} \code{GETFIELD_PRIMITIVE} e t b f \\
{\bf Desc.:} Generated after thread t reads primitive-typed instance field f of object b at program point e.

\item
Getfield reference \\
{\bf Event:} \code{GETFIELD_REFERENCE} e t b f o \\
{\bf Desc.:} Generated after thread t reads object o from reference-typed instance field f of object b at program point e.

\item
Putfield primitive \\
{\bf Event:} \code{PUTFIELD_PRIMITIVE} e t b f \\
{\bf Desc.:} Generated after thread t writes primitive-typed instance field f of object b at program point e.

\item
Putfield reference \\
{\bf Event:} \code{PUTFIELD_REFERENCE} e t b f o \\
{\bf Desc.:} Generated after thread t writes object o to reference-typed instance field f of object b at program point e.

\item
Aload primitive \\
{\bf Event:} \code{ALOAD_PRIMITIVE} e t b i \\
{\bf Desc.:} Generated after thread t reads the primitive-typed element at index i of array object b at program point e.

\item
Aload reference \\
{\bf Event:} \code{ALOAD_REFERENCE} e t b i o \\
{\bf Desc.:} Generated after thread t reads object o from the reference-typed element at index i of array object b at program point e.

\item
Astore primitive \\
{\bf Event:} \code{ASTORE_PRIMITIVE} e t b i \\
{\bf Desc.:} Generated after thread t writes the primitive-typed element at index i of array object b at program point e.

\item
Astore reference \\
{\bf Event:} \code{ASTORE_REFERENCE} e t b i o \\
{\bf Desc.:} Generated after thread t writes object o to the reference-typed element at index i of array object b at program point e.

\item
Method call \\
{\bf Event:} \code{BEF_METHOD_CALL} i t o \\
{\bf Desc.:} Generated before thread t executes the method invocation statement at program point i with this argument as object o. \\
{\bf Event:} \code{AFT_METHOD_CALL} i t o \\
{\bf Desc.:} Generated after thread t executes the method invocation statement at program point i with this argument as object o. \\
{\bf Note:} These events are not generated before/after constructor calls; use \code{BEF_NEW} and \code{AFT_NEW} events for that purpose.

\item
Thread start call \\
{\bf Event:} \code{THREAD_START} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{start()} method of \code{java.lang.Thread} at program point i and spawns a thread o.

\item
Thread join call \\
{\bf Event:} \code{THREAD_JOIN} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread o.

\item
Lock acquire \\
{\bf Event:} \code{ACQUIRE_LOCK} l t o \\
{\bf Desc.:} Generated after thread t executes a statement of the form monitorenter o or enters a method synchronized on o at program point l.

\item
Lock release \\
{\bf Event:} \code{RELEASE_LOCK} r t o \\
{\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.

\item
Thread wait call \\
{\bf Event:} \code{WAIT} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{wait()} method of \code{java.lang.Object} at program point i on object o.

\item
Thread notify call \\
{\bf Event:} \code{NOTIFY} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point i on object o.
\end{enumerate}