### 2010-12-18 main/doc/program_representation.tex

\section{Program representation}
\label{sec:program-representation}

TODO