### 2010-12-18 main/doc/user_guide.tex

\documentclass{article}
\usepackage{hyperlatex}
\usepackage{hyperref}

\texonly{\oddsidemargin=0in}
\texonly{\evensidemargin=0in}
\texonly{\textwidth=6.0in}
\texonly{\textheight=8.0in}

\htmlonly{\newcommand\url[1]{\xlink{#1}{#1}}}
\htmlonly{\newcommand\code[1]{{\tt #1}}}
\texonly{\renewcommand\code[1]{\path{#1}}}

\htmltitle{Chord}
\htmladdress{mhn@cs.stanford.edu}

\title{Chord: A Static and Dynamic Program Analysis Framework for Java}
\author{Mayur Naik}
\date{\today}
\begin{document}
\maketitle

\input{whatis_chord}
\texonly{\newpage}
\input{download_and_installation}
\texonly{\newpage}
\input{program_setup}
\texonly{\newpage}
\input{building_scope}
\texonly{\newpage}
\input{running_analysis}
\texonly{\newpage}
\input{writing_analysis}
\texonly{\newpage}
\input{writing_dynamic_analysis}
\texonly{\newpage}
\input{instr_events}
\texonly{\newpage}
\input{writing_datalog_analysis}
\texonly{\newpage}
\input{tuning_datalog_analysis}
\texonly{\newpage}
\input{bdd_representation}
\texonly{\newpage}
\input{program_representation}
\texonly{\newpage}
\input{pointsto}
\texonly{\newpage}
\input{datarace}
\texonly{\newpage}
\input{deadlock}
\texonly{\newpage}
\input{chord_sysprops}
\texonly{\newpage}
\input{acknowledgments}

\end{document}