### 2011-03-12 main/doc/acknowledgments.tex
more documentation

\section{Acknowledgments}
\label{sec:acknowledgments}

Chord would not be possible without the following open-source software:

\begin{itemize}
\item
\xlink{Joeq}{http://joeq.sourceforge.net/}, a Java compiler framework
\item
\xlink{Javassist}{http://www.csg.is.titech.ac.jp/~chiba/javassist/},
a Java bytecode manipulation tool
\item
\xlink{bddbddb}{http://bddbddb.sourceforge.net/}, a BDD-based Datalog solver
\end{itemize}

Chord additionally relies on the following open-source tools and libraries:
\begin{itemize}
\item
\xlink{Ant-Contrib}{http://ant-contrib.sourceforge.net/},
a collection of useful Ant tasks  
\item
\xlink{BuDDy}{http://buddy.sourceforge.net/}, a BDD library
\item
\xlink{GNU Trove}{http://trove4j.sourceforge.net/},
a primitive collections library for Java
\item
\xlink{Java2HTML}{http://www.java2html.com/} and
\xlink{Java2Html}{http://www.java2html.de/}, Java-to-HTML tools
\item
\xlink{Saxon}{http://saxon.sourceforge.net/}, an XSLT processor
\end{itemize}

Chord was supported in part by grants from the National Science Foundation,
an equipment grant from Intel, and a Microsoft fellowship during 2005-2007.