### 2011-03-12 main/doc/running_chord.tex

\section{Running Chord}
\label{sec:running-chord}

Running Chord requires a JVM with JDK 5 or higher.
There are two equivalent ways to run Chord.
One way is to run the following command in the \code{main/} directory:

\begin{verbatim}
    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
\end{verbatim}

The above requires Apache Ant to be installed.  The alternative, which
does not require Apache Ant, is to invoke Chord directly as follows:

\begin{verbatim}
    prompt> java -cp <...>/chord.jar -D<key1>=<value1> ... -D<keyN>=<valueN> chord.project.Boot
\end{verbatim}

where \code{<...>} denotes the absolute or relative path of the directory containing file \code{chord.jar};
that directory is also expected to contain any other binaries in Chord's installation
(e.g., \code{libbuddy.so} and \code{libchord_instr_agent.so}).