### 2011-03-12 main/doc/troubleshooting.tex

\section{Troubleshooting}
\label{sec:troubleshooting}

This section describes known common problems with running \Chord\ and how to fix or work around them.
We assume that the following command was run in \Chord's {\tt main/} directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...> <sys-props> run
\end{verbatim}
\end{quote}

where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
of the form {\tt -D<key>=<value>} (see Section \ref{sec:chord-sysprops} for an exhaustive listing of all
properties recognized by \Chord).

Abnormal termination of the above command typically produces output of the form:

\begin{quote}
\begin{verbatim}
Buildfile: build.xml

props-file:

pathconvert:

run:
     [java] java version "1.6.0"
     [java] ... // JVM version details omitted
     [java] 
     [java] Redirecting stdout to file: <...>/chord_output/log.txt
     [java] Redirecting stderr to file: <...>/chord_output/log.txt
     [java] Java Result: 1

BUILD FAILED
<...>/main/build.xml:325: Chord terminated abnormally; see above file(s) for details.
\end{verbatim}
\end{quote}

The first step is to inspect the file(s) reported above to which stdout and stderr were redirected.
We next list common errors and how to fix or work around them.

\begin{enumerate}
\item
\end{enumerate}