### 2011-03-14 main/doc/predefined_analyses.tex

\chapter{Predefined Analyses}
\label{chap:predefined-analyses}

This chapter describes various predefined analyses in Chord.

\section{Static Datarace Analysis}

To run Chord's static datarace analysis, run the following
command in Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
\end{verbatim}

where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
meaning of these properties.

Directory \code{main/examples/datarace_test/} provides a toy Java
program on which you can run the datarace analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{examples/datarace_test/}.  Upon successful completion, the
following files should be produced in directory
\code{examples/datarace_test/chord_output/}:

\begin{itemize}
\item
File \code{dataraces_by_fld.html}, listing all dataraces grouped by
the field on which they occur; all dataraces on the same instance
field or the same static field are listed in the same group, and so
are all dataraces on array elements.
\item
File \code{dataraces_by_obj.html}, listing all dataraces grouped by
the abstract object on whose field they occur; dataraces on all static
fields are listed in the same group, and so are dataraces on different
instance fields of the same abstract object.
\end{itemize}

\section{Static Deadlock Analysis}

To run Chord's static deadlock analysis,  run the following
command in Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
\end{verbatim}

where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Chapter \ref{chap:properties} for the
meaning of these properties.

Directory \code{main/examples/deadlock_test/} provides a toy Java
program on which you can run the deadlock analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{examples/deadlock_test/}.  Upon successful completion, the
file \code{deadlocks.html} should be produced in directory
\code{examples/deadlock_test/chord_output/}.

\section{May-Alias and Call-Graph Analyses}

TODO