### 2011-03-17 main/doc/predefined_analyses.tex
add material describing predefined points-to analyses

\xname{predefined_analyses}
\chapter{Predefined Analyses}
\label{chap:predefined-analyses}

This chapter describes various predefined analyses in Chord.

\section{Static Datarace Analysis}

To run Chord's static datarace analysis, run the following
command in Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=datarace-java run
\end{verbatim}

where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:program-props} for the
meaning of these properties.

Directory \code{main/examples/datarace_test/} provides a toy Java
program on which one can run the datarace analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{examples/datarace_test/}.  Upon successful completion, the
following files should be produced in directory
\code{examples/datarace_test/chord_output/}:

\begin{itemize}
\item
File \code{dataraces_by_fld.html}, listing all dataraces grouped by
the field on which they occur; all dataraces on the same instance
field or the same static field are listed in the same group, and so
are all dataraces on array elements.
\item
File \code{dataraces_by_obj.html}, listing all dataraces grouped by
the abstract object on whose field they occur; dataraces on all static
fields are listed in the same group, and so are dataraces on different
instance fields of the same abstract object.
\end{itemize}

\section{Static Deadlock Analysis}

To run Chord's static deadlock analysis,  run the following
command in Chord's {\tt main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> -Dchord.run.analyses=deadlock-java run
\end{verbatim}

where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:program-props} for the
meaning of these properties.

Directory \code{main/examples/deadlock_test/} provides a toy Java
program on which one can run the deadlock analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{examples/deadlock_test/}.  Upon successful completion, the
file \code{deadlocks.html} should be produced in directory
\code{examples/deadlock_test/chord_output/}.

\section{Points-to and Call-Graph Analyses}

Chord offers several choices for computing points-to and call graphs for programs.
The simplest of these is a context-insensitive analysis. We describe this first,
because understanding the concepts behind it will help you understand the more
sophisticated context-sensitive analysis.

In general, context-sensitive approaches are more precise but run more slowly;
the higher the degree of precision, the slower the analysis.

\subsection{Context-insensitive points-to}

The context-insenstive points-to is implemented in \code{analyses/alias/cipa-0cfa.dlog}.
The analysis treats all objects allocated at the same site as interchangeable.
Put another way, each allocation site defines a unique abstract object.
Domain H is the set of all allocation sites and therefore also of all abstract objects.
The analysis produces several output relations.

\begin{description}
\item[VH] contains tuples \texttt{(v,h)} if variable \texttt{v} can point to abstract object \texttt{h}.
\item[MM] contains tuples \texttt{(m1,m2)} if method \texttt{m1} might call \texttt{m2}.
\item[IM] contains tuples \texttt{(i,m)} if invocation \texttt{i} might call \texttt{m}.
\item[reachableM] contains the set of potentially-reachable methods.
\item[reachableT] contains the set of potentially-reachable types.
\item[reachableI] contains the set of potentially-reachable invocation statements.
\item[rootM] contains the set of potentially-reachable methods without a caller (such as \texttt{main} and static initializers).
\item[FH] contains tuples \texttt{(f,h)} if static field \texttt{f} can point to abstract object \texttt{h}.
\item[HFH] contains tuples \texttt{(b,f,h)} if instance field \texttt{f} of abstract object \texttt{b} can point to abstract object \texttt{h}.
\end{description}

\subsection{Context-sensitive}

In a context-sensitive analysis, there is no longer just one abstract object per site.
Rather, the set of objects a reference can point to depends on the \textit{context} of
the method call containing the reference.
Whereas a context-insensitive analysis talks about the domain of methods and the domain
of allocation sites (M and H), a context-sensitive analysis talks about the domain of contexts, labelled C.
Domain C refers to both abstract objects and also to abstract calling contexts. (As we will see, these
are merged for a reason.)

Chord has several context-sensitive analyses, but they all expose the same relations.

\begin{description}
\item[C] domain containing all abstract contexts.
\item[CC] contains tuples \texttt{(c,c2)} such that \texttt{c2} is all but the last element of context \texttt{c}.
\item[CH] contains tuples \texttt{(c,h)} such that object allocation site \texttt{h} is the last element of abstract context \texttt{c}.
\item[CI] contains tuples \texttt{(c,i)} such that call site \texttt{i} is the last element of abstract context \texttt{c}.
\item[CVC] contains tuples \texttt{(c,v,o)} such that local \texttt{v} might point to object \texttt{o} in context \texttt{c} of its declaring method.  \\
Note that both \texttt{o} and \texttt{c} are elements of domain C.
\item[CFC] contains tuples \texttt{(o1,f,o2)} such that instance field \texttt{f} of object \texttt{o1} might point to object \texttt{o2}.
\item[FC] contains tuples \texttt{(f,o)} such that static field \texttt{f} may point to object \texttt{o}.
\item[CICM] contains tuples \texttt{(c,i,c2,m)} if invocation \texttt{i} in context \texttt{c} can reach method 2 (in context \texttt{c2}) .
\item[rootCM] contains tuples \texttt{(c,m)} such that method \texttt{m} is an entry method in context \texttt{c}.
\item[reachableCM] contains tuples \texttt{(c,m)} such that method \texttt{m} can be called in context \texttt{c}.
\end{description}

{\bf Under Construction: need to explain object-sensitive vs context-sensitive, plus how to invoke.}