### 2011-03-18 extra/sliver_pldi11/abstractions.tex

\Section{abstractionExamples}{Abstractions}

Recall from \refsec{preliminaries} that we separate the Datalog program that
computes the query from the abstraction.  In \refsec{klimited}, we described
the Datalog programs that computed the queries for our three clients based on
$\infty$-object-sensitivity.  We now describe the abstractions that use to
further abstract this analysis.

The basic abstraction (\refsec{klimitedAbstraction}) corresponds exactly to
$k$-object-sensitivity.  We then present two orthogonal variants: one that in
addition limits the repetition of allocation sites (\refsec{barelyAbstraction})
and one that further abstracts allocation sites using type information
(\refsec{classAbstraction}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{klimitedAbstraction}{$k$-limited abstraction}

The {\em $k$-limited abstraction}
essentially partitions chains based on their length $k$ prefix.
We first setup some notation.
%Let $\dom$ be a set of values (e.g., $\dom = \{ 0, 1 \}$).
For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and let
let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
For two chains $c_1, c_2$, let $c_1+c_2$ denote concatenation.
For a chain $c$, let $c*$ denote the set of all chains
with prefix $c$; formally:
\begin{align}
c* \eqdef \{ c' \times \H^* : c'[1..|c|] = c \}.
\end{align}
%It is important that $c*$ is a set.

For an integer truncation level $k \ge 0$, define the abstraction $\klimabs_k$ as follows:
\begin{align}
\label{eqn:klimitedAbstraction}
\klimabs_k(c) \eqdef
\begin{cases}
\{c\}     & \text{if $|c| < k$} \\
c[1..k]*  & \text{if $|c| \ge k$}. \\
\end{cases}
\end{align}
If the concrete chain $c$ is shorter than length $k$,
we map it to the singleton set $\{ c \}$;
otherwise, we map it to the set of chains that share the first $k$ elements.
It is easy to verify that $\klimabs_k$ as defined is a valid abstraction under \refdef{abstraction}.
\reffig{repeatingExample} shows an example.

\begin{figure}
\[
\arraycolsep=3.5pt
\begin{array}{llllllll}
\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
\{\vec{00}\}   &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \{\vec{11}\} \\
\vec{000}*     & \vec{001}* & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* & \vec{110}* & \vec{111}*
\end{array}
\]
\caption{\label{fig:repeatingExample} For the $k$-limited abstraction
with $\dom = \{ 0, 1 \}$ and $k = 3$, the equivalence classes defined by $\klimabs_k$.
}
\end{figure}

Traditionally, $k$-limited analyses
simply use a list of sites ``$c$'' to represent an abstract value,
not making a distinction between $\{c\}$ and $c*$;
however when we do pruning, this distinction is crucial.
Suppose that $k=3$.
Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
The problem comes when we prune.
If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
This is unacceptable because pruning tuples should not change the abstraction.
Actually, the situation is worse, because (i) in fact no pruning
(in the sense that some chains are ignored)
would happen and (ii) the resulting analysis would be much more imprecise because
``$[\,]$'' is always be present for representing the context of the main method,
but this is a prefix of any chain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{barelyAbstraction}{Barely-repeating $k$-limited abstraction}

When we applied the $k$-limited abstraction empirically,
we noticed that a major reason why it did not scale was
the seemingly unnecessary combinatorial explosion associated chains formed
by cycling endlessly through the same allocation sites.
For $k$-CFA, this repetition corresponds to recursion.
For $k$-object-sensitivity, this corresponds to recursive allocation, as
illustrated in \reffig{recursion}.\footnote{
Incidentally, the example in the figure also gives an interesting example where
$k$-object-sensitivity for any finite $k$ (no matter how large) is less precise
than $\infty$-object-sensitivity.}
We therefore wish to define an abstraction that not only truncates at $k$ but also truncates
a chain when it starts repeating.

\begin{figure}
\begin{center}
\begin{minipage}{2in}
\begin{verbatim}
class A {
  f() {
0:  v = new A
    if (*) return v
    else return v.f()
  }
}
\end{verbatim}
\end{minipage}
\begin{minipage}{1in}
\begin{verbatim}
1:  x1 = new A
2:  x2 = new A
    y1 = x1.f()
    y2 = x2.f()
\end{verbatim}
\end{minipage}
\end{center}
\caption{\label{fig:recursion} An example illustrating the repetition of allocation sites.
The points-to set of {\tt y1} using $\infty$-object-sensitivity
is $\{\vec{01},\vec{001},\vec{0001},\dots\}$ (any positive number of zeros followed by a 1),
and the points-to set of {\tt y2} is $\{\vec{02},\vec{002},\vec{0002},\dots\}$.
These two sets are disjoint, but for any finite $k$, no matter how large,
$k$-object-sensitivity would conclude that both point to $\bzero_k*$, where $\bzero_k$ is a chain of $k$ zeros.
}
\end{figure}

For a sequence $c$, we say $c$ is {\em non-repeating} if all its elements are distinct.
We say $c$ is {\em barely-repeating} if $c$ excluding the last element ($c[1..|c|-1]$)
is {\em non-repeating}.  Note that a barely-repeating includes the non-repeating case.
Let $\delta(c)$ be the length of the longest prefix that is barely-repeating:
\begin{align}
\label{eqn:barelyLen}
\distinct(c) \eqdef \max_{m' : c[1..m'] \text{ is barely-repeating}} m'.
\end{align}
For example, $\distinct(\vec{10010}) = 3$ because $\vec{100}$ is barely-repeating, but $\vec{1001}$ is not.

Then we define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
\begin{align}
\label{eqn:barelyAbstraction}
\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\delta(c)\}}(c),
\end{align}
\reffig{barelyExample} shows an example of $\klimdabs_k$.
It is not automatic that $\klimabs_k$ is a valid abstraction,
but we can check that it is:
\begin{proposition}
\label{prop:barelyAbstraction}
The function $\klimdabs_k$ defined in \refeqn{barelyAbstraction} is a valid abstraction
(\refdef{abstraction}).
\end{proposition}
\begin{proof}
We consider two cases: (i) for $\{c\} \in \range(\klimdabs_k)$,
we have $\klimdabs_k(c) = \{c\}$; and (ii) for any $c* \in \range(\klimdabs_k)$,
either $|c| = k$ or $c$ is barely-repeating; in either case,
we can check that any extension $c' \in c*$ will have $\klimabs_k(c') = c*$.
\end{proof}
Remark: one might wonder why we defined the abstraction using the barely-repeating criterion
as opposed to the simpler non-repeating criterion.
It turns out that using the latter in \refeqn{barelyLen} would not result in a valid abstraction.
If $\klimdabs_k$ were defined using the non-repeating criterion,
then $\klimdabs_3(\vec{00}) = \vec{0}*$.
But for $\vec{01} \in \vec{0}*$, we have $\klimdabs_3(\vec{01}) = \{\vec{01}\} \neq \vec{0}*$.

\begin{figure}
\[
\begin{array}{lllllll}
\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
\vec{00}*      &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \vec{11}* \\
               &            & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* &
\end{array}
\]
\caption{\label{fig:barelyExample} For the barely-repeating $k$-limited abstraction
for $\dom = \{0,1\}$ and $k = 3$, we show the equivalence classes under $\klimdabs_k$.
Compare this with the classes for the $k$-limited abstraction (\reffig{repeatingExample}).
Note that, for example,
$\vec{000}*$ and $\vec{001}*$ are collapsed into $\vec{00}*$ since $\vec{000}$ and $\vec{001}$ are not barely-repeating,
but $\vec{00}$ is.
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{classAbstraction}{Class-based abstraction}

We now introduce an abstraction that we will use in the pre-pruning step of the \PR\ algorithm.
We start by defining an equivalence relation over allocation sites $\dom$,
represented by a function $\hclass : \dom \mapsto \sP(\dom)$
mapping each allocation site $h \in \dom$ to its its equivalence class.
Given such such an $\hclass$, we could then extend it sequences by elementwise application:
\begin{align}
\hclass(c) = \hclass(c[1]) \times \cdots \times \hclass(c[|c|]), \quad c \in \dom^*.
\end{align}

To construct $\hclass$, we consider two sources of type information associated
with an allocation site, motivated by \cite{smaragdakis11context}:
\begin{align}
\typeIs(h)  &= \text{declaring type of allocation site $h$} \label{eqn:typeIs} \\
\typeHas(h) &= \text{type of class containing allocation site $h$} \label{eqn:typeHas}
\end{align}
Using these two functions, we can construct three equivalence relations,
$\hclassIs$, $\hclassHas$, and $\hclassIsHas$ as follows:
\begin{align}
\hclass_f(h) = \{ h' : f(h) = f(h') \}, \label{eqn:hclass}
\end{align}
with $f \in \{ \typeIs, \typeHas, \typeIs \times \typeHas \}$.

% Complementary
Now we have three options for $\hclass$ which are complementary to the
$k$-limited abstraction: $\hclass$ abstracts a chain by abstracting each site
uniformly; $\pi_k$ keeps all sites up to a length $k$ prefix precisely but
abstracts away the rest completely.  This complementarity is desirable
for pre-pruning.
However, $\hclass$ is not coarser than $\klimabs_k$, and therefore cannot
be used directly in the \PR\ algorithm.

However, we can compose $\hclass$ and $\klimabs_k$ to yield another abstraction which will be coarser than $\klimabs_k$.
But in what order should we compose?
We must take care because the composition of two equivalence relations is not necessarily an equivalence relation.
The following proposition shows which orderings are valid:
\begin{proposition}
The functions
(i) $\klimabs_k \circ \hclass$ and (ii) $\hclass \circ \klimabs_k$
are valid abstractions (see \refdef{abstraction}) and equivalent;
(iii) $\klimdabs_k \circ \hclass$ is also valid,
but (iv) $\hclass \circ \klimdabs_k$ is not.
\end{proposition}
\begin{proof}
For each of these four composed functions, each set $s$ in the range of the function
must be either of the form $s = w_1 \times \cdots \times w_m$ for $m < k$ (case 1)
or $s = w_1 \times \cdots \times w_m \times \dom^*$ for some $m \le k$ (case 2),
where $w_i \in \range(\hclass)$ for each $i = 1, \dots, m$.

For (i) and (ii), it is straightforward to check
$(\klimabs_k \circ \hclass)(c) = (\hclass \circ \klimabs_k)(c) = s$ for each $c \in s$.
Intuitively, the truncation ($\klimabs_k$) and coarsening ($\tau$) operate
independently and can be interchanged.

For (iii) and (iv), the two dimensions do not act independently;
the amount of truncation depends on the amount of coarsening:
the coarser $\hclass$ is, the more truncation one might need to limit repetitions.
Showing that $\klimdabs_k \circ \hclass$ is valid
proceeds in a similar manner to \refprop{barelyAbstraction}.
If $s$ falls under case 1, note that no $c \in s$ is repeating
because the $w_i$'s must be disjoint; therefore $\klimdabs_k(\tau(c)) = s$.
If $s$ falls under case 2, note that for any $c[1..m] \in s$ must be barely-repeating but any longer prefix is not,
and therefore, $\klimdabs_k(\tau(c)) = s$.

To show that (iv) is not an abstraction,
consider the following counterexample:
let $\dom = \{ 0,1,2 \}$,
and define $\hclass(v) = \dom$ for all $v \in \dom$
(there is one equivalence class).
Consider two elements $\vec{01}$ and $\vec{00}$ under the composed abstraction:
For $\vec{01}$, we have $\klimdabs_3(\vec{01}) = \{ \vec{01 }\}$,
so $\hclass(\klimdabs_3(\vec{01})) = \dom^2$;
for $\vec{00}$, we have $\klimdabs_3(\vec{00}) = \vec{00}*$,
so $\hclass(\klimdabs_3(\vec{00})) = \dom^2 \times \dom^*$.
But $\dom^2 \subsetneq \dom^2 \times \dom^*$ (notably,
the two sets are neither equal nor disjoint),
so $\hclass \circ \klimdabs_3$ does not define an equivalence relation.
\end{proof}

In light of this result, we will use the valid abstractions $\cklimabsk$ and $\cklimdabsk$,
which work by first applying $\hclass$ and then applying $\klimabs_k$ or $\klimdabs_k$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\Subsection{computation}{Computation}
%
%In order to use the PR algorithm (\reffig{pseudocode}),
%we need to implement the various steps efficiently.
%
%We now present techniques for computing 
%
%\paragraph{Pre-pruning step}
%
%Computing $\cklimdabsk$
%
%\paragraph{Pruning step}
%
%However, need to compute $\ext$.
%
%\paragraph{Refinement step}
%
%We use $\alpha_t = \klimdabs_{t+1}$
%and $\beta_t = \klimdabs_{t+1}$
%and we set $\hclass(h)$ to be the dynamic type of allocation site $h$.
%
%Computing $\klimdabs_{k+1}$
%
%Suppose we have we have a current $A_k \subset \range(\klimdabs_k)$.
%For $c \in \dom^*$.
%$\alpha_{k+1}(\{c\}) = $
%Given $c$, we consider