### 2011-03-18 extra/sliver_pldi11/conclusion.tex

\Section{conclusion}{Conclusion}

We have introduced pruning as a general way of scaling up static analyses
written as Datalog programs.  The key idea is to run an analysis using a coarse
abstraction, only keeping input tuples deemed relevant for a finer abstraction.
Theoretically, we showed that pruning is both sound and complete (our analysis
is valid and we lose no precision).  Empirically, we showed that our pruning
approach enables us to scale up analyses based on $k$-object-sensitivity much
more than previous approaches.