### 2011-03-18 extra/sliver_pldi11/introduction.tex

\Section{introduction}{Introduction}

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
supporting abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction increases.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis,
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
The general theme of all these approaches is that the
complexity of the abstraction is increased only in parts deemed useful using
some feedback from a client query.

% Pruning idea: general
In this paper, we use {\em pruning}, a different approach from
the above selected refinement techniques for making static analyses scale.
Our approach works on any analysis that can be expressed as a Datalog program.
A Datalog program takes a set of input tuples (determined by the abstraction)
and derives new tuples via a set of inference rules; the answer to a client
query corresponds to whether a designated tuple can be derived.  The key idea
is to identify input tuples which are provably irrelevant for deriving the
query tuple and prune these away.  Selected refinement techniques
attempt to keep the set of input tuples small by not refining some of them; we
keep the set small by removing some of them entirely.

% Soundness
Pruning can be a dangerous affair though.  With selected refinement, at any
point we are still performing static analysis with respect to an abstraction
and therefore have soundness guarantees.  However, once we start pruning input
tuples, we are no longer running a valid static analysis and thus do not
automatically inherit soundness guarantees.  Nonetheless, we prove that our
method is {\em sound with respect to the client}.

% Completeness
While soundness is trivial for selected refinement but requires some argument for pruning,
the situation is reversed for {\em completeness}, that is, the guarantee that
an analysis is as precise as some target.  For selected refinement,
it is difficult to argue that the chosen coarser abstraction is as precise as if we
had refined everything.  However, with pruning, we are working directly with
the more complex abstraction, and thus automatically inherit its completeness
guarantees.  By removing input tuples, we can only prove more queries.

% Refinement algorithm
%To capitalize on the idea of {\em pruning}, we propose an algorithm, which we
%call Refinement-Pruning (\PR).  The idea is to start with a coarse abstraction
%and prune the input tuples before moving to a more expensive abstraction.

% Slivers
We apply our pruning technique to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects on the heap are abstracted into a chain of
allocation sites.  Pruning in this setting corresponds to maintaining a subset
of these chains.  Although the main contribution of the paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise because we are pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction
than before due to pruning.  Second, we introduce an abstraction which limits
chains to length $k$ and also truncates to avoid having too many repeating
sites; this is an effective way to increase $k$ further (possible now due to
pruning) without getting bogged down by recursion.  Third, we show that an
abstraction that replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having an accurate pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
algorithm is able to go well beyond $k=10$.