### 2011-03-18 extra/sliver_pldi11/preliminaries.tex

\Section{preliminaries}{Preliminaries}

The pruning approach that we present in this paper works on Datalog, a general
language that has been used recently to declaratively express static analyses \cite{Whaley2007,bravenboer}.  Normally, one thinks of
the Datalog program as encapsulating the abstraction, but it will be useful for us
to decouple the two.\footnote{The reason is that we want to make theoretical
statements comparing the same Datalog program across different
abstractions.}  We first define Datalog (\refsec{datalog}), which describes the
computation of a query (think more in the direction of concrete semantics).
Then, we focus on the abstraction (\refsec{abstractions}), which only interacts
with the Datalog program via the input tuples.
Throughout this section, we will use \reffig{graphExample} as a running example.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{center} {\bf Graph Example} \end{center}
{\bf Input relations}:
\[
\begin{array}{ll}
\edge(g,i,j) & \defn{edge from node $i$ to node $j$ in graph $g$} \\
\head(c,i)   & \defn{first element of array $c$ is $i$} \\
\ext(i,c,c') & \defn{$i$ prepended to $c$ yields $c'$: $c' = [i]+c$} \\
\end{array}
\]
{\bf Rules}:
\[
\begin{array}{lcl}
\path(g,[\initNode]). & \\
\path(g,c')           &\dlogUpdate& \path(g,c), \head(c,i), \\
                      &           & \edge(g, i, j), \ext(j, c, c'). \\
\common(g_1,g_2,i)    &\dlogUpdate& \path(g_1,c), \path(g_2,c), \head(c,i). \\
\end{array}
\]
{\bf Query tuple}: $\xo = \common(\Ga,\Gb,3)$. \\
\\
{\bf Constants}: $\sC = \{ \Ga, \Gb, 0, 1, 2, 3, \vec{0}, \vek{01}, \dots \}$.
\\
\\
\includegraphics[scale=0.5]{figures/graphExample}
\caption{\label{fig:graphExample}
A simple example illustrating Datalog:
Suppose we have two graphs $\Ga$ and $\Gb$ defined on the same set of nodes $\{0,1,2,3\}$,
and we want to compute the query tuple $\common(\Ga,\Gb,3)$,
asking whether the two graphs have a common path from node $0$ to node $3$.
Given the input tuples encoding the graph,
the Datalog program computes a set of derived tuples from the rules.
In this case, the absence of $\common(\Ga,\Gb,3)$ means the query is false.
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{datalog}{Datalog}

A {\em Datalog program} consists of
a set of {\em constants} $\sC$ (e.g., $\initNode,\vec{03} \in \sC$),
a set of {\em variables} $\sV$ (e.g., $i, j \in \sV$), and
a set of {\em relations} $\sR$ (e.g., $\edge \in \sR$).

A {\em term} $t$ consists of a relation $t.r \in \sR$ and a list of arguments $t.\ba$,
where each argument $t.a_i$ is either a variable or a constant,
(that is, $t.a_i \in \sV \cup \sC$) for $i = 1, \dots, |t.\ba|$.
We will write a term in any of the following three equivalent ways:
\begin{align}
t \quad\equiv\quad t.r(t.\ba) \quad\equiv\quad t.r(t.a_1, \dots, t.a_{|t.\ba|}).
\end{align}
For example, $\ext(j,c,c')$ is a term.
We call a term whose arguments are all constants a {\em tuple}
(e.g., $\ext(\initNode, [\,], \vec{\initNode})$).
Note that the tuple includes the relation as well as the arguments.
We let $\xo$ denote a designated {\em query tuple} (e.g., $\common()$),
whose truth value we want to determine.

Let $\sZ$ denote the set of rules,
where each {\em rule} $z \in \sZ$ consists of a target term $z.t$ and a set of source terms $z.\bs$.
We write a rule if $z.t = t$ and $z.\bs = \{ s_1, \dots, s_k \}$ as
\begin{align}
t \dlogUpdate s_1, \dots, s_n.
\end{align}

An {\em assignment} is a function $f : \sV \mapsto \sC$ which maps variables to
constants.  To simplify notation later, we extend an assignment $f$ so that it
can be applied (i) to constants ($f(c) = c$ for $c \in \sC$), and (ii) to terms by
replacing the variables in the term with constants ($f(t) = t.r(f(t.a_1), \dots, f(t.a_{|t.\ba|}))$).

\begin{figure}[ht]
%{\bf Notation:} \\
\[
\begin{array}{ll}
\xo                           & \defn{designated query tuple} \\
\sC                           & \defn{set of concrete values} \\
\bD(X)                        & \defn{derivations using input tuples $X$} \\
\bP(X) \subset X              & \defn{set of relevant input tuples} \\
\alpha : \sC \mapsto \sP(\sC) & \defn{abstraction, maps to equivalence class} \\
A_k \subset \sP(\sC)          & \defn{abstract input tuples after $k$ iterations} \\
\tilde A_k = \bP(A_k)         & \defn{relevant abstract input tuples} \\
\end{array}
\]
\caption{\label{fig:notation} Notation.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\Subsection{derivations}{Derivations}
\paragraph{Derivations}

Intuitively, a Datalog program takes a set of input tuples and derives new
tuples.  To formalize this computation, we define the notation of a derivation.

A {\em derivation} (of the query $\xo$) with respect to a set of input tuples
$X$ is a sequence $\bx = (x_1, \dots, x_n)$ such that
\begin{enumerate}
\item [(i)] for each $i = 1, \dots, n$, we have $x_i \in X$; or
there exists a set of indices $J$ such that $(J,i)$ satisfies the following conditions:
$j < i$ for each $j \in J$, and
there is a rule $z \in \sZ$ and an assignment $f$
such that $f(z.t) = x_i$ and
$\{ x_j : j \in J \} = \{ f(s) : s \in z.\bs \}$.
\item [(ii)] $x_n = \xo$; and
\item [(iii)] 
for each $j = 1, \dots, n\!-\!1$,
there exists $J$ such that $j \in J$ and an index $i$
such that $(J,i)$ satisfies the conditions in (i).
\end{enumerate}
Define $\bD(X)$ to be the set of all derivations with respect to the input tuples $X$.

% Explain
Condition (i) says that each tuple in a derivation
should either be given as an input tuple ($x_i \in X$)
or be the result of some rule $z \in \sZ$.
Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
Condition (iii) says that the derivation of $x_n$ is minimal; each tuple was
used in some way; as we will see later, this condition is important because it allow us to
define the set of relevant tuples for pruning.

A Datalog program computes $\bD(X)$.
We say that the query $\xo$ is false (proven) if and only if $\bD(X) = \emptyset$.
Although the query is the ultimate quantity of interest,
the Datalog program can be used to provide more information that is useful for pruning.
Specifically, we define $\bP(X)$ to be the subset of the input tuples $X$
which were used in some derivation:
\begin{align}
\label{eqn:Pdef}
\bP(X) \eqdef \{ x \in X : x \in \bx \in \bD(X) \}.
\end{align}
We call $\bP(X)$ the set of {\em relevant input tuples}.  As we will see later,
any tuple not in this set can be safely pruned.
In fact, $\bP(X)$ also tells us whether the query is true or false.
In particular, $\bD(X) = \emptyset$ if and only if $\bP(X) = \emptyset$.
This equivalence suggests that proving and pruning are intimately related;
in some sense, proving the query is just pruning away the query tuple.
In the remainder of the paper, we will make heavy use of $\bP$ as the principal
proving/pruning operation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Computation}

We can compute $\bP(X)$ in Datalog by using the Datalog program transformation technique described in 
\cite{liang11minimal}; this allows us to use off-the-shelf Datalog solvers which have already been optimized.
Specifically, we define a set of new relations $\sR' = \{ r' : r \in \sR \}$.
For a term $t = t.r(t.\ba)$ we let $t' = t.r'(t.\ba)$ be the term that using the corresponding new relation.
We then add the following new Datalog rules:
\begin{align}
\xo' &\dlogUpdate \xo. \label{eqn:revBase} \\
s'   &\dlogUpdate z.t', z.\bs. \label{eqn:indBase}
\end{align}
The key is that a tuple $x'$ can be derived by the new Datalog program if and only if $x \in \bP(X)$.
These two rules construct $\bP(X)$ recursively:
The base case \refeqn{revBase} states that the query tuple $\xo \in \bP(X)$.
The recursive case \refeqn{indBase} states that if $x \in \bP(X)$ and a rule $z$ (with some assignment $f$)
was used to produce $x$, then for every source term $s \in z.\bs$ of that rule,
we also have $f(s) \in \bP(X)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{abstractions}{Abstractions}

Given a Datalog program, we define an {\em abstraction} as a equivalence relation over
the constants $\sC$ (concrete values).  In particular, we represent it as the projection
function which maps each element to its equivalence class (its abstract value):
\begin{definition}
\label{def:abstraction}
An abstraction is a function $\alpha : \sC \to \sP(\sC)$
such that for each set $s \in \range(\alpha)$,
we have $\alpha(c) = s$ for all $c \in s$.
\end{definition}
We use the natural partial order on abstractions,
where $\alpha_1 \preceq \alpha_2$ iff $\alpha_1(c) \supset \alpha_2(c)$ for all $c$,
that is, $\alpha_2$ is finer than $\alpha_1$.

% Example
In the context of our example (\reffig{graphExample}),
we could define an abstraction
that maps a sequence $c$ to the set of sequences that have the same first element as $c$
(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
Applied to a tuple, we get
\begin{align}
& \alpha(\ext(1, [0], [10])) = \\
& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
\end{align}
This abstraction is a simplified version of the abstractions we use in our
$k$-limited analyses (see \refsec{abstractions}).

% Extend to sets
We extend $\alpha$ to sets of constants (which we also call abstract values):
\begin{align}
\alpha(s) = \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
\end{align}
which returns a set of abstract values.
This allows us to naturally define compositions of abstractions
by flattening sets of abstract values:
\begin{align}
(\alpha_1 \circ \alpha_2)(c) \eqdef \cup_{s \in \alpha_1(\alpha_2(c))} s.
\end{align}
An important property of equivalence relations is
that if $\alpha_1 \preceq \alpha_2$, then $\alpha_1 \circ \alpha_2 = \alpha_1$
(applying a finer abstraction first has no impact).
Note that in general for arbitrary $\alpha_1,\alpha_2$,
the composition $\alpha_1 \circ \alpha_2$ is not an abstraction;
it is something that we must check when we consider compositions in
\refsec{abstractions}.

% Extend to tuples
We also extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$:
\begin{align}
\alpha(x) &= x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
\alpha(X) &= \{ \alpha(x) : x \in X \}.
\end{align}
Here, $\alpha(x)$ is an abstract tuple (one where the arguments are abstract values)
and $\alpha(X)$ is a set of abstract tuples.
Finally, we extend $\alpha$ to abstract tuples $b$ and sets of abstract tuples $B$:
\begin{align}
\alpha(b) &= \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
\alpha(B) &= \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
\end{align}
\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the possible abstract values,
yielding a set of abstract tuples;
\refeqn{alphaB} aggregates these sets of abstract tuples.

Given an abstraction $\alpha$,
we can run an abstract version of the Datalog program to compute an abstract answer to the query tuple.
We do this by applying the abstraction to the concrete input tuples $X$,
to produce a set of abstract input tuples $\alpha(X)$.
We then can feed these tuples into the Datalog program,
which is oblivious to whether the tuples are abstract or concrete;
conceptually, the Datalog program now operates on the constants $\sP(\sC)$ rather than $\sC$.
\reffig{graphDerivation} shows an example of performing this computation on the
graph example from \reffig{graphExample}.

We say the query is proven by $\alpha$ if $\bP(\alpha(X)) = \emptyset$,
and it is proven only if the query is actually false ($\bP(X) = \emptyset$).
This is the standard soundness property, summarized below (see \refapp{proofs} for the proof):
\begin{proposition}[Abstraction is sound]
\label{prop:soundness}
Let $\alpha$ be an abstraction and let $X$ be any set of input tuples.
If $\bP(\alpha(X)) = \emptyset$ (the query is false abstractly),
then $\bP(X) = \emptyset$ (the query is false concretely).
\end{proposition}

\FigStar{figures/graphDerivation}{0.3}{graphDerivation}{
Computation of $\bP(X)$ on the graph example from \reffig{graphExample}
under an abstraction which maps a path onto the set of paths with the same first element
(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
Each abstract tuple is derived by a rule whose source terms corresponding to the incoming edges.
Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;
ones which are not are pruned ($X\backslash\bP(X)$, shown in red).
}