### 2011-03-18 extra/sliver_pldi11/related.tex

\Section{related}{Related Work}

% Larger space of abstractions
There is a wealth of literature which attempts to scale static analyses without
sacrificing precision.  One general theme is to work with a flexible family of
abstractions, which in principle allows us to conform to the needs of the
client.  Milanova et al.~\cite{kobj,MilanovaRountevRyder2005} also consider
abstractions, where each local variable can be independently treated context
sensitively or insensitively, and different $k$ values can be chosen for
different allocation sites.  Lhot\'{a}k and
Hendren~\cite{LhotakHendren2006,LhotakHendren2008} present Paddle, a
parametrized framework for BDD-based, $k$-limited alias analysis.
\cite{smaragdakis11context} scale up $k$-object-sensitivity one more level by
using types rather than allocation sites.  However, which parts of the
abstraction should be more refined is largely left up to the user.

% Client-driven
Client-driven approaches use feedback from a client query to determine what
parts of an abstraction to refine.
Plevyak and Chien~\cite{PlevyakChien1994} use a refinement-based algorithm for
type inference, where context sensitivity is driven by detecting type conflict.
Guyer and Lin~\cite{GuyerLin2003} present a pointer analysis for C
which detects loss of precision (e.g., at merge points) and introduce context sensitivity.
Our method for determining relevant input tuples is similar in spirit but more general.

Perhaps the closest to this work is the part of Liang et al.~\cite{liang11minimal}, which
uses the same technique to compute the set of relevant tuples.  However, what
is done with this information is quite different.  \cite{liang11minimal} merely
stops refining the irrelevant sites whereas we actually prune all irrelevant
tuples, thereby exploiting this information fuller.  As we saw in
\refsec{experiments}, this difference had major ramifications.
%An importance difference is that we are not

%Similar to iterative refinement client-driven approaches is lazy abstraction
%\cite{henzinger02lazy,mcmillan06lazy} from model checking, which

% Demand-driven
Demand-driven analyses \cite{HeintzeTardieu2001,ZhengRugina2008} do not refine
the abstraction but rather compute an analysis on an existing
abstraction faster.  Some approaches such as \cite{SridharanBodik2006} do both.
Our \PR\ algorithm has a flavor of the client-driven approach, in that we do
refine our abstraction, but also of the demand-driven approach, in that we do
not perform a full computation (in particular, ignoring tuples which were
pruned).

%In contrast to client-driven analyses, which compute an exhaustive solution of
%varying precision, demand-driven approaches compute a partial solution of fixed
%precision.  Heintze and Tardieu's demand-driven alias analysis for
%C~\cite{HeintzeTardieu2001} performs a provably optimal computation to
%determine the points-to sets of variables queried by the client. Their analysis
%is applied to call-graph construction in the presence of function pointers. A
%more recent alias analysis developed by Zheng and Rugina~\cite{ZhengRugina2008}
%uses a demand-driven algorithm that is capable of answering alias queries
%without constructing points-to sets.

% Model checking
%Lazy abstraction \cite{henzinger02lazy} is an abstraction refinement technique
%for model checking by which one starts with a small set of predicates and tries
%to prove a query.  If the proof fails, an abstract counterexample (derivation
%in our terminology) is generated, and predicates are refined on this path.
%One difference is that they refine one predicate at a time.
%
%This work has been extended to interpolants \cite{mcmillan06lazy}, which
%provide a richer (and as they show, more effective) class of abstraction
%predicates.

%Our family of abstractions is still $\sA = \{ \klimabs_k : k = 0, 1 \}$.
%Let $K : \H \to \{ 0, 1, 2, \dots \}$ be a function mapping each site to a $k$ value.
%In \cite{liang11minimal}, we considered the abstraction family:
%\begin{align}
%\klimabs_K(c) = \klimabs_{K(c[1])}(c).
%\end{align}

% Dynamic - lower bound
Pruning has been implemented in various settings.
\cite{vipindeep05pruning} use dynamic analysis to prune down set of paths and
then focus a static analysis on these paths.
\cite{spoon04ddp} use pruning for type inference in functional languages,
where pruning is simply a heuristic which shortcuts a search algorithm.
As a result, pruning can hurt precision.
One advantage of our pruning approach is that it comes with strong soundness
and completeness guarantees.

%This is another fruitful but orthogonal way to prune.  The intuition is that if
%a query was observed to be true dynamically, there is no point trying to prove
%it statically.
%For example, we can use the dynamic analysis framework of
%\cite{liang10abstraction} to obtain lower bounds, predicates which are
%certainly true.  If we observe dynamically observe that a variable $v$ points
%to an object with abstraction $a$, the static analysis need not even attempt to
%prove that $v$ does not point to $a$.