### 2011-03-20 extra/sliver_pldi11/abstract.tex

\begin{abstract}

Many static analyses do not scale as they are made more precise.  For example,
increasing the amount of context sensitivity in a $k$-limited pointer analysis causes
the number of contexts to grow exponentially.  Iterative refinement techniques
can mitigate this growth by starting with a coarse abstraction and
only refining parts of the abstraction that are deemed relevant with respect to a client.

In this paper, we introduce a new technique called {\em pruning} that uses this client
feedback to greatly reduce the complexity of an analysis.  The pruned analysis is no
longer a valid abstraction, but we prove that we do not forfeit correctness and
that our analysis is still sound with respect to a given client. 
For $k$-limited pointer analysis, our approach amounts to keeping track of a set
of context prefix patterns.  By iteratively refining and pruning these
patterns, we show that our analysis can greatly increase scalability.

\end{abstract}