### 2011-03-20 extra/sliver_pldi11/abstract.tex

\begin{abstract}

Many static analyses do not scale as they are made more precise.  For example,
increasing the amount of context sensitivity in a $k$-limited pointer analysis causes
the number of contexts to grow exponentially with $k$.  Iterative refinement techniques
can mitigate this growth by starting with a coarse abstraction and
only refining parts of the abstraction that are deemed relevant with respect to a given client.

In this paper, we introduce a new technique called {\em pruning} that uses client
feedback in a different way.  The basic idea is to use coarse
abstractions to prune away parts of the program analysis deemed irrelevant for proving a
client query, and then using finer abstractions on the sliced program analysis.
%In contrast to existing iterative refinement techniques,
%our analyses do not correspond to abstractions on the original program but rather one on a sliced program.
%We prove that we do not forfeit correctness and
%that our analysis is still sound with respect to a given client. 
For a $k$-limited pointer analysis, this approach amounts to adaptively refining and pruning a set
of prefix patterns representing the contexts relevant for the client.
By pruning, we are able to scale up to much more expensive abstractions than before.
We also prove that the pruned analysis is both sound and complete, that is, it yields the same results
as an analysis that uses a more expensive abstraction directly without pruning.

\end{abstract}