### 2011-03-21 extra/sliver_pldi11/abstractions.tex

\Section{abstractionExamples}{Abstractions}

%Recall from \refsec{preliminaries} that we separate the Datalog program that
%computes the query from the abstraction.  In \refsec{klimited}, we described
%the Datalog programs that computed the queries for our three clients based on
%$\infty$-object-sensitivity.  We now describe the abstractions that we use to
%further abstract this analysis.

We have already defined the $k$-limited abstraction, which corresponds to
$k$-object-sensitivity.  We now present two orthogonal variants of this basic abstraction: one that
additionally limits the repetition of allocation sites (\refsec{barelyAbstraction})
and one that further abstracts allocation sites using type information
(\refsec{classAbstraction}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{barelyAbstraction}{Barely-repeating $k$-limited abstraction}

When we applied the $k$-limited abstraction in practice,
we noticed empirically that a major reason why it did not scale was
the seemingly unnecessary combinatorial explosion associated with chains formed
by cycling endlessly through the same allocation sites.
For $k$-CFA, this repetition corresponds to recursion.
For $k$-object-sensitivity, this corresponds to recursive allocation, as
illustrated in \reffig{recursion}.\footnote{
Incidentally, the example in the figure also gives an interesting example where
$k$-object-sensitivity for any finite $k$ (no matter how large) is less precise
than $\infty$-object-sensitivity.}
We therefore wish to define an abstraction that not only truncates chains at length $k$ but also truncates
a chain when it starts repeating.

\begin{figure}
\begin{center}
\begin{minipage}{2in}
\begin{verbatim}
class A {
  f() {
0:  v = new A
    if (*) return v
    else return v.f()
  }
}
\end{verbatim}
\end{minipage}
\begin{minipage}{1in}
\begin{verbatim}
1:  x1 = new A
2:  x2 = new A
    y1 = x1.f()
    y2 = x2.f()
\end{verbatim}
\end{minipage}
\end{center}
\caption{\label{fig:recursion} An example illustrating the repetition of allocation sites.
The points-to set of {\tt y1} using $\infty$-object-sensitivity
is $\{\vec{01},\vec{001},\vec{0001},\dots\}$ (any positive number of zeros followed by a 1),
and the points-to set of {\tt y2} is $\{\vec{02},\vec{002},\vec{0002},\dots\}$.
While these two sets are disjoint, if we use a $k$-limited abstraction for any finite $k$,
we would conclude erroneously that both variables might point to $\bzero_k*$, where $\bzero_k$ is a chain of $k$ zeros.
Incidentally, this demonstrates an intrinsic limitation of the $k$-object-sensitivity abstraction.
Using the barely-repeating $k$-limited abstraction, we can increase $k$ while avoiding chains longer than $\vec{00}$
since $\vec{00}$ is barely-repeating.
This results in computational savings, and in this case, in no loss in precision.
}
\end{figure}

For a sequence $c$, we say $c$ is {\em non-repeating} if all its elements are distinct.
We say $c$ is {\em barely-repeating} if (i) $c$ excluding the last element ($c[1..|c|-1]$)
is {\em non-repeating} and (ii) the last element of $c$ is repeated earlier in $c$.
%We say $c$ is {\em barely-repeating} if $c$ excluding the last element ($c[1..|c|-1]$) is {\em non-repeating}.
%Note that a barely-repeating includes the non-repeating case.
Let $\distinct(c)$ be the length of the longest prefix of $c$ that is barely-repeating, if it exists, and $\infty$
otherwise:
\begin{align}
\label{eqn:barelyLen}
\distinct(c) \eqdef
\begin{cases} 
\max_{m' : c[1..m'] \text{ is barely-repeating}} m' & \text{if $m'$ exists}, \\
\infty & \text{otherwise}.
\end{cases}
\end{align}
%\begin{align}
%\label{eqn:barelyLen}
%\distinct(c) \eqdef \max_{m' : c[1..m'] \text{ is barely-repeating}} m'
%\end{align}
For example, $\distinct(\vec{10010}) = 3$ because $\vec{100}$ is barely-repeating, but $\vec{1001}$ is not.

We now define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
\begin{align}
\label{eqn:barelyAbstraction}
\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\distinct(c)\}}(c),
\end{align}
\reffig{barelyExample} shows an example of $\klimdabs_k$.
We show that $\klimdabs_k$ is a valid abstraction:
\begin{proposition}
\label{prop:barelyAbstraction}
The function $\klimdabs_k$ defined in \refeqn{barelyAbstraction} is a valid abstraction
(\refdef{abstraction}).
\end{proposition}
\begin{proof}
We consider two cases: (i) for $\{c\} \in \range(\klimdabs_k)$,
we have $\klimdabs_k(c) = \{c\}$; and (ii) for any $c* \in \range(\klimdabs_k)$,
either $|c| = k$ or $c$ is barely-repeating; in either case,
it is easy to see that any extension $c' \in c*$ will have $\klimdabs_k(c') = c*$.
\end{proof}
Remark: one might wonder why we defined the abstraction using the barely-repeating criterion
as opposed to the simpler non-repeating criterion.
It turns out that using the latter in \refeqn{barelyLen} would not result in a valid abstraction.
If $\klimdabs_k$ were defined using the non-repeating criterion,
then $\klimdabs_3(\vec{00}) = \vec{0}*$.
But for $\vec{01} \in \vec{0}*$, we have $\klimdabs_3(\vec{01}) = \{\vec{01}\} \neq \vec{0}*$.

\begin{figure}
\[
\begin{array}{lllllll}
\{\vec{0}\}    &            &              &            & \{\vec{1}\} \\
\vec{00}*      &            & \{\vec{01}\} &            & \{\vec{10}\} &            & \vec{11}* \\
               &            & \vec{010}*   & \vec{011}* & \vec{100}*   & \vec{101}* &
\end{array}
\]
\caption{\label{fig:barelyExample} For the barely-repeating $k$-limited abstraction
for $\dom = \{0,1\}$ and $k = 3$, we show the equivalence classes under $\klimdabs_k$.
Compare this with the classes for the $k$-limited abstraction (\reffig{repeatingExample}).
Note that, for example,
$\vec{000}*$ and $\vec{001}*$ are collapsed into $\vec{00}*$ since $\vec{000}$ and $\vec{001}$ are not barely-repeating,
but $\vec{00}$ is.
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{classAbstraction}{Type-based abstraction}

We now introduce an abstraction that we will use in the pre-pruning step of the Prune-Refine algorithm.
We start by defining an equivalence relation over allocation sites $\dom$,
represented by a function $\hclass : \dom \mapsto \sP(\dom)$
mapping each allocation site $h \in \dom$ to its equivalence class.
In the graph example, we might have
$\hclass(0) = \hclass(1) = \{ 0,1 \}$
and $\hclass(2) = \hclass(3) = \{ 2,3 \}$.

Given such a $\hclass$, we extend it to sequences by taking the cross product over elementwise applications:
\begin{align}
\hclass(c) = \hclass(c[1]) \times \cdots \times \hclass(c[|c|]), \quad c \in \dom^*.
\end{align}
In the running example, $\hclass([02]) = \{ [02], [03], [12], [13] \}$.

To construct $\hclass$ for $k$-limited pointer analysis, we consider using two sources of type information associated
with an allocation site, motivated by \cite{smaragdakis11context}:
\begin{align}
\typeIs(h)  &= \text{declaring type of allocation site $h$} \label{eqn:typeIs} \\
\typeHas(h) &= \text{type of class containing allocation site $h$} \label{eqn:typeHas}
\end{align}
Using these two functions, we can construct three equivalence relations,
$\hclassIs$, $\hclassHas$, and $\hclassIsHas$ as follows:
\begin{align}
\hclass_f(h) = \{ h' : f(h) = f(h') \}, \label{eqn:hclass}
\end{align}
for $f \in \{ \typeIs, \typeHas, \typeIs \times \typeHas \}$.

% Complementary
Now we have three choices for $\hclass$: one that uses the declaring type ($\hclassIs$),
one that uses the type of the containing class ($\hclassHas$), and one that uses both ($\hclassIsHas$).
Recall that all three are complementary to the
$k$-limited abstraction $\pi_k$: Specifically, $\hclass$ abstracts a chain by abstracting each site
uniformly, whereas $\pi_k$ performs no abstraction on the first $k$ sites, but performs a total abstraction
on the rest of the chain.  Recall that this complementarity is desirable for effective pre-pruning.

Since $\hclass$ is not coarser than $\klimabs_k$, we cannot
use it directly in the Prune-Refine algorithm.
We must compose $\hclass$ with $\klimabs_k$ or $\klimdabs_k$ to yield another
abstraction which is coarser than $\klimabs_k$ or $\klimdabs_k$,
respectively.  But in what order should we compose?
We must be careful because the composition of two abstractions is not necessarily an abstraction.
Fortunately, the following proposition shows which compositions are valid:
\begin{proposition}
The functions
(i) $\klimabs_k \circ \hclass$ and (ii) $\hclass \circ \klimabs_k$
are valid abstractions (see \refdef{abstraction}) and equivalent;
(iii) $\klimdabs_k \circ \hclass$ is also valid,
but (iv) $\hclass \circ \klimdabs_k$ is not.
\end{proposition}
\begin{proof}
For each of these four composed functions, each set $s$ in the range of the function
must be either of the form $s = w_1 \times \cdots \times w_m$ for $m < k$ (case 1)
or $s = w_1 \times \cdots \times w_m \times \dom^*$ for some $m \le k$ (case 2),
where $w_i \in \range(\hclass)$ for each $i = 1, \dots, m$.

For (i) and (ii), it is straightforward to check that
$(\klimabs_k \circ \hclass)(c) = (\hclass \circ \klimabs_k)(c) = s$ for each $c \in s$.
Intuitively, the truncation ($\klimabs_k$) and coarsening ($\tau$) operate
independently and can be interchanged.

For (iii) and (iv), the two dimensions do not act independently;
the amount of truncation depends on the amount of coarsening:
the coarser $\hclass$ is, the more truncation one might need to limit repetitions.
Showing that $\klimdabs_k \circ \hclass$ is valid
proceeds in a similar manner to \refprop{barelyAbstraction}.
If $s$ falls under case 1, note that no $c \in s$ is repeating
because the $w_i$'s must be disjoint; therefore $\klimdabs_k(\tau(c)) = s$.
If $s$ falls under case 2, note that for any $c[1..m] \in s$ must be barely-repeating but any longer prefix is not,
and therefore, $\klimdabs_k(\tau(c)) = s$.

To show that (iv) is not an abstraction,
consider the following counterexample:
let $\dom = \{ 0,1,2 \}$,
and define $\hclass(h) = \dom$ for all $h \in \dom$
(there is one equivalence class).
Consider applying $\hclass \circ \klimdabs_3$ to two elements $\vec{01}$ and $\vec{00}$:
For $\vec{01}$, we have $\klimdabs_3(\vec{01}) = \{ \vec{01 }\}$,
so $\hclass(\klimdabs_3(\vec{01})) = \dom^2$;
for $\vec{00}$, we have $\klimdabs_3(\vec{00}) = \vec{00}*$,
so $\hclass(\klimdabs_3(\vec{00})) = \dom^2 \times \dom^*$.
But $\dom^2 \subsetneq \dom^2 \times \dom^*$ (notably,
the two sets are neither equal nor disjoint),
so $\hclass \circ \klimdabs_3$ does not define a valid abstraction.
\end{proof}

\noindent 
In light of this result, we will use the valid abstractions $\cklimabsk$ and $\cklimdabsk$,
which work by first applying the type-based abstraction $\hclass$ and then applying $\klimabs_k$ or $\klimdabs_k$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\Subsection{computation}{Computation}
%
%In order to use the PR algorithm (\reffig{pseudocode}),
%we need to implement the various steps efficiently.
%
%We now present techniques for computing 
%
%\paragraph{Pre-pruning step}
%
%Computing $\cklimdabsk$
%
%\paragraph{Pruning step}
%
%However, need to compute $\ext$.
%
%\paragraph{Refinement step}
%
%We use $\alpha_t = \klimdabs_{t+1}$
%and $\beta_t = \klimdabs_{t+1}$
%and we set $\hclass(h)$ to be the dynamic type of allocation site $h$.
%
%Computing $\klimdabs_{k+1}$
%
%Suppose we have we have a current $A_k \subset \range(\klimdabs_k)$.
%For $c \in \dom^*$.
%$\alpha_{k+1}(\{c\}) = $
%Given $c$, we consider