### 2011-03-21 extra/sliver_pldi11/conclusion.tex

\Section{conclusion}{Conclusion}

We have introduced pruning as a general technique for scaling up static analyses
written in Datalog.  The basic idea is to run an analysis using a coarse
abstraction, only keeping input tuples deemed relevant, and then using a finer abstraction on the remaining tuples.
Theoretically, we showed that pruning is both sound and complete (our analysis
is valid and we lose no precision).  Empirically, we showed that pruning
enables us to scale up analyses based on $k$-object-sensitivity much
more than previous approaches.

\acks
We thank Mooly Sagiv and Hongseok Yang for discussion and useful feedback.
We also thank the anonymous reviewers for their insightful comments.