### 2011-03-27 main/doc/user_guide.tex

\documentclass[openany]{book}
\usepackage{hyperlatex}
\usepackage{hyperref}
\texonly{
	\usepackage{graphicx}
	\usepackage{framed}
}

\texonly{
% put borders around figures
	%\usepackage{float}
	%\floatstyle{boxed} 
	%\restylefloat{figure}
% set dimensions of page/text
	\parindent=0in
	\parskip=10pt
	\oddsidemargin=-0.3in
	\evensidemargin=-0.3in
	\textwidth=7in
	\textheight=8.0in
% redefine certain commands
	\renewcommand\code[1]{\path{#1}}
	\renewcommand\xlink[2]{\href{#2}{#1}}
}

\htmlonly{
	\newcommand\code[1]{{\tt #1}}
	\newcommand\url[1]{\xlink{#1}{#1}}
	\newenvironment{framed}{\xmlattributes*{table}{border=1 frame=box rules=none cellpadding=8}\begin{tabular}{l}}{\end{tabular}}
}

\newcommand\bh{{\bf h}}
\newcommand\bi{{\bf i}}
\newcommand\bo{{\bf o}}
\newcommand\bt{{\bf t}}
\newcommand\bm{{\bf m}}
\newcommand\be{{\bf e}}
\newcommand\bg{{\bf f}}
\newcommand\bb{{\bf b}}
\newcommand\bl{{\bf l}}
\newcommand\bp{{\bf p}}
\newcommand\br{{\bf r}}
\newcommand\bw{{\bf w}}

\setcounter{htmldepth}{2}
\setcounter{htmlautomenu}{3}

\htmltitle{Chord}
\htmladdress{mhn@cs.stanford.edu}

\title{Chord: A Versatile Platform for Program Analysis}
\author{Mayur Naik}
\date{\today}
\begin{document}
\maketitle

\input{whatis_chord}
\input{getting_started}
\input{example}
\input{architecture}
\input{properties}
\input{predefined_analyses}
\input{analyses}
\input{dynamic_analysis}
\input{datalog_analysis}
\input{computing_scope}
\input{program_representation}
\input{acknowledgments}

\end{document}