### 2011-04-04 main/doc/getting_started.tex
documentation

\xname{getting_started}
\chapter{Getting Started}
\label{chap:getting-started}

This chapter describes how to download, install, and run Chord.
Section \ref{sec:downloading-binaries} describes how to obtain
pre-built binaries of Chord.  Section \ref{sec:downloading-sources}
describes how to obtain the source code of Chord and Section
\ref{sec:downloading-sources} explains how to build it.  Finally,
Section \ref{sec:running-chord} describes how to run Chord.

\section{Downloading Binaries}
\label{sec:downloading-binaries}

To obtain Chord's pre-built binaries, download and uncompress file
\xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
It includes the following files:

\begin{enumerate}
\item
\code{chord.jar}, which contains the class files of Chord and of
libraries used by Chord.
\item
\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you
can keep one of these files depending upon whether you intend to run
Chord on Linux, Windows/Cygwin, or MacOS, respectively.  These files
are needed only if you want BDD library BuDDy to be used when the
BDD-based Datalog solver bddbddb in Chord runs analyses written in
Datalog.
\item
\code{libchord_instr_agent.so}: this file is needed only if you want
the JVMTI-based bytecode instrumentation agent to be used when Chord
runs dynamic analyses.
\end{enumerate}

Novice users can ignore items (2) and (3) until they become more
familiar with Chord.  The binaries mentioned in items (2) and (3)
might not be compatible with your machine, in which case you can
either forgo using them (with hardly any noticeable difference in
functionality), or you can download the sources (see Section
\ref{sec:downloading-sources}) and build them yourself (see Section
\ref{sec:compiling-sources}).

\section{Downloading Source Code}
\label{sec:downloading-sources}

To obtain Chord's source code, download and uncompress the following
files:

\begin{itemize}
\item
Mandatory: file
\xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz},
which contains Chord's source code and jars of libraries used by
Chord.
\item
Optional: file
\xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz},
which contains the source code of libraries used by Chord (e.g., joeq,
javassist, bddbddb, etc.)
\end{itemize}

Alternatively, you can obtain the latest development snapshot from the
SVN repository by running the following command:

\begin{framed}
\begin{verbatim}
svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}
\end{framed}

Instead of checking out the entire \code{trunk/}, which contains
several sub-directories, you can check out specific sub-directories:

\begin{itemize}
\item
\code{main/} contains Chord's source code and jars of libraries used by Chord.
\item
\code{libsrc/} contains the source code of libraries used by Chord
(e.g., joeq, javassist, bddbddb, etc.).
\item
\code{test/} contains Chord's regression tests.
\item
many more; these might eventually move under \code{main/}.
\end{itemize}

Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz}
mentioned above are essentially stable releases of the \code{main/}
and \code{libsrc/} directories, respectively.

\section{Compiling the Source Code}
\label{sec:compiling-sources}

Compiling Chord's source code requires the following software:

\begin{itemize}
\item
JVM for Java 5 or higher, e.g.
\xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java build tool.
\end{itemize}

Chord's main directory contains a file named {\tt build.xml} which is
interpreted by Apache Ant.  To see the various possible targets,
simply run command ``{\tt ant}" in that directory.

To compile Chord, run command ``{\tt ant compile}" in the same directory.
This will compile Chord's Java sources from \code{src/} to class
files in \code{classes/}, as well as build a jar file
\code{chord.jar} that contains these class files as well as those
in the jars of libraries that are used by Chord and are
provided under \code{lib/} (e.g., \code{joeq.jar},
\code{javassist.jar}, \code{bddbddb.jar}, etc.).  Additionally:

\begin{itemize}
\item

If system property \code{chord.use.buddy} is set to \code{true}, then
the C source code of BDD library
\xlink{BuDDy}{http://buddy.sourceforge.net/} from directory \code{bdd/}
will be compiled to a shared library named
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS; this library is used by BDD-based
Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} in
Chord for running analyses written in Datalog.

\item

If system property \code{chord.use.jvmti} is set to \code{true}, then
the C++ source code of the JVMTI-based bytecode instrumentation agent
from directory \code{agent/} will be compiled to a shared library named
\code{libchord_instr_agent.so} on all architectures;
this agent is used in Chord for computing analysis scope dynamically
and for running dynamic analyses.
\end{itemize}

Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are
defined in file \code{chord.properties} in Chord's main directory.
The default value of both these properties is \code{false}.  If you set
either of them to \code{true}, then you will also need a utility like
GNU Make (to run the \code{Makefile}'s in directories \code{bdd/} and
\code{agent/}) and a C++ compiler (to build the above shared libraries).

\section{Running Chord}
\label{sec:running-chord}

Running Chord requires a JVM supporting Java 5 or higher.
There are two equivalent ways to run Chord.
One way, which is available only in the source installation of Chord, is to run the
following command in Chord's main directory:

\begin{framed}
\begin{verbatim}
ant -D<key1>=<val1> ... -D<keyN>=<valN> run
\end{verbatim}
\end{framed}

The above requires Apache Ant (a Java build tool) to be installed on
your machine.  The alternative, which does not require Apache Ant and is
available in both the source and binary installations of Chord, is
to run the following command:

\begin{framed}
\begin{verbatim}
java -cp <...>/chord.jar -D<key1>=<val1> ... -D<keyN>=<valN> chord.project.Boot
\end{verbatim}
\end{framed}

where \code{<...>} denotes the absolute or relative path of the
directory containing file \code{chord.jar}; that directory is also
expected to contain any shared libraries in Chord's installation (e.g.,
\code{libbuddy.so} and \code{libchord_instr_agent.so}).

Each ``\code{-D<key>=<val>}" argument above sets the system property
named \code{<key>} to the value denoted by \code{<val>}.  The only
way to specify inputs to Chord is via system properties; there is no
command-line argument processing.  Chapter \ref{chap:properties}
describes all system properties recognized by Chord.