### 2011-04-15 main/doc/organization.tex
more documentation

\xname{organization}
\chapter{Organization of this Guide}
\label{chap:organization}

Chord is a program analysis platform that is both {\it stand-alone}, in that it
provides many standard analyses for users to run, and {\it extensible}, in that
it allows users to write their own analyses, possibly atop the provided
analyses.
As a result, Chord has two kinds of users: {\it end-users}, who only wish to run predefined
analyses, and {\it developers}, who additionally wish to write and run their
own analyses.

For convenience, this user guide consists of two parts: a
\xlink{guide for end-users}{enduser\_guide.html} and a
\xlink{guide for developers}{developer\_guide.html}.  Unlike end-users,
developers need to understand Chord's source code and API, and code written by
them is executed as part of a Chord run.  Hence, the guide for end-users
concerns how to run Chord, and the guide for developers concerns how to extend
Chord.