### 2011-04-15 main/doc/predefined.tex

\xname{predefined}
\chapter{Predefined Analyses}
\label{chap:predefined}

Chord provides many standard analyses.
This chapter first explains how to run any such analysis and then provides
descriptions of various predefined analyses.

\section{Running an Analysis}
\label{sec:running-predefined}

Each predefined analysis in Chord has a unique name that can be used to run the
analysis from the command-line.  The following command runs the analysis named
\code{<ANALYSIS_NAME>} on the program specified by directory \code{<WORK_DIR>}
(see Chapter \ref{chap:setup} for how to setup a program):

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=<ANALYSIS_NAME> run
\end{verbatim}
\end{framed}

For instance, the following command runs a basic may-alias and call-graph
analysis (called 0CFA) provided in Chord:

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
\end{verbatim}
\end{framed}

This instructs Chord to run the analysis named \code{cipa-0cfa-dlog},
which is defined in file \code{main/src/chord/analyses/alias/cipa_0cfa.dlog}.

The output of the above command is of the form:

\begin{framed}
{\small
\begin{verbatim}
Buildfile: build.xml

run:
     [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
     [java] ENTER: cipa-0cfa-dlog
     [java] ENTER: T
     [java] ENTER: RTA
     [java] Iteration: 0
     [java] Iteration: 1
     [java] Iteration: 2
     [java] LEAVE: RTA
     [java] SAVING dom T size: 1386
     [java] LEAVE: T
     [java] ENTER: F
     [java] SAVING dom F size: 4120
     [java] LEAVE: F
     ...
     [java] ENTER: MputStatFldInst
     [java] SAVING rel MputStatFldInst size: 739
     [java] LEAVE: MputStatFldInst
     [java] ENTER: statIM
     [java] SAVING rel statIM size: 3359
     [java] LEAVE: statIM
     [java] Starting command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
     [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
     [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
     [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
     [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
     [java] Finished command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
     [java] LEAVE: cipa-0cfa-dlog
     [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
     [java] Total time: 00:00:27:671 hh:mm:ss:ms

BUILD SUCCESSFUL
Total time: 28 seconds
\end{verbatim}
}
\end{framed}

Each analysis in Chord is written modularly, independent of other analyses,
along with lightweight annotations specifying the inputs and outputs of the
analysis.  Chord's runtime automatically computes producer-consumer
relationships between analyses (e.g., determines which analysis produces as
output a result that is needed as input by another analysis).  Before running a
desired analysis (such as 0CFA in the above example), Chord recursively runs
other analyses until the inputs to the desired analysis have been computed; it
finally runs the desired analysis to produce the outputs of that analysis.

The 0CFA analysis consumes and produces multiple {\it program relations}.  The
consumed program relations include \code{MputStatFldInst} and \code{statIM},
each of which is produced by a separate imperative analysis with the
corresponding name, and the produced program relations include \code{VH},
\code{FH}, \code{HFH}, and \code{IM}.  We next briefly discuss these relations.

The program relations consumed by the 0CFA analysis contain basic program facts.
For instance, \code{MputStatFldInst} is a relation containing each tuple
($m$,$f$,$v$) such that method $m$ in the input Java program contains a
putstatic instruction of the form ``$f$ = $v$", while \code{statIM} is a
relation containing each tuple ($i$,$m$) such that $m$ is the target method of
invokestatic instruction $i$.

The program relations produced by the 0CFA analysis represent points-to
information and the call graph of the input Java program as computed by the
analysis.  Specifically, relations \code{VH}, \code{FH}, and \code{HFH}
represent points-to information for local variables, static fields, and instance
fields, respectively, while relation \code{IM} represents the call graph,
namely, it contain each tuple ($i$,$m$) such that $m$ is a possible target
method of call site $i$.

Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called
{\it program domains} \code{M}, \code{F}, \code{I}, and \code{V}, respectively.
A program domain is a set of entities of a certain kind in the input Java
program.  For instance, \code{M} is the domain representing the set of all
methods in the input Java program, \code{F} is the domain representing the set
of all fields, \code{I} is the domain representing the set of all call sites in
methods in \code{M}, and \code{V} is the domain representing the set of all
local variables of reference type in methods in \code{M}.  Each of these domains
is produced by a separate Java analysis with the corresponding name.  Notice
that the analyses producing these domains run upfront because these domains are
consumed by the analyses that produce relations such as \code{MputStatFldInst}
and \code{statIM}, which in turn are consumed by the desired 0CFA analysis.

During execution, Chord dumps intermediate and final results to files in the
directory specified by property \code{chord.out.dir}, whose default value is
\code{[chord.work.dir]/chord_output/} and typically does not need to be changed
by users.  For the above example, this directory is
\code{<WORK_DIR>/chord_output/}.

The verbosity of Chord's output above is controlled by property
\code{chord.verbose}, whose default value is 1.  At verbosity level 0, the above
command produces less voluminous output of the form:

\begin{framed}
{\small
\begin{verbatim}
Buildfile: build.xml

run:
     [java] Chord run initiated at: Mar 13, 2011 10:35:01 PM
     [java] Chord run completed at: Mar 13, 2011 10:35:28 PM
     [java] Total time: 00:00:26:297 hh:mm:ss:ms

BUILD SUCCESSFUL
Total time: 26 seconds
\end{verbatim}
}
\end{framed}

\section{Points-to and Call-Graph Analyses}

Chord offers several choices for computing points-to and call-graph information
of Java programs.  In each of these choices, the points-to and call-graph
information is computed simultaneously (called ``on-the-fly call-graph
construction" in the literature in contrast to ``ahead-of-time call-graph
construction" in which the call graph is computed first followed by points-to
information).  On-the-fly approaches are more precise because, in a dynamically
dispatching language like Java, as more points-to facts are discovered, more
(dynamically dispatched) methods are deemed reachable, thereby growing the call
graph; the code in these newly added methods in turn results in more points-to
facts.

Flow-insensitive analysis computes a single abstract heap whereas flow-sensitive
analysis computes per-program-point abstract heaps.  Context-insensitive
analysis analyzes each method at most once (i.e. in a single abstract context),
whereas context-sensitive analyses potentially analyze each method multiple
times, in different abstract contexts.  Thus, flow- and context-sensitive
analyses are more precise but less scalable than flow- and context-insensitive
analyses, respectively.

Flow-sensitive analysis does not offer much precision over flow-insensitive
analysis in practice, especially in the absence of strong updates and in the
presence of SSA (Static Single Assignment form), a program representation that
renders a flow-insensitive analysis almost as precise as a flow-sensitive
analysis.  Since analyses in Chord currently perform only weak updates, and
since they all operate on an SSA form of the input Java program by default, the
rest of this section focuses only on flow-insensitive analysis, which is the
predominant kind of points-to/call-graph analysis in Chord.

We describe context-insensitive analysis first because understanding the
concepts behind it will help understand the more sophisticated context-sensitive
analyses.  We first recall some relevant program domains:

\texonly{\newpage}

\begin{itemize}
\item M is the domain of all methods.
\item I is the domain of all method call sites.
\item F is the domain of all (instance and static) fields.
\item V is the domain of all local variables of reference type.
\item H is the domain of all object allocation sites.
\end{itemize}

\subsection{Context-Insensitive Analysis}

The context-insensitive points-to/call-graph analysis treats each object
allocation site as a separate abstract memory location; in other words, it can
distinguish objects created at different sites but not those created at the same
site.  Additionally, it is field-sensitive, in that it distinguishes between
different instance fields of the same object, but array-insensitive, in that it
cannot distinguish between different elements of the same array; all array
elements are modeled using a distinguished hypothetical instance field (that has
index 0 in domain F).

To run this analysis, run the following command:

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
\end{verbatim}
\end{framed}

where \code{<WORK_DIR>} is a directory
containing a file named \code{chord.properties} that defines
properties \code{chord.main.class} and \code{chord.class.path} specifying
the main class and the application classpath, respectively, of the program
to be analyzed.

This analysis outputs the following relations:

%is implemented in \code{chord/analyses/alias/cipa-0cfa.dlog}.

\begin{itemize}
\item {\it Call-graph information:}
\begin{itemize}
\item
rootM subset M contains the set of entry methods that may be reachable; this
includes the program's main method as well as each static initializer method
that may be reachable from the main method.
\item
reachableM subset M contains the set of methods that may be reachable from the
program's main method.
\item
IM subset (I $\times$ M) contains tuples \texttt{(i,m)} such that call site
\texttt{i} may call method \texttt{m}.
%\item
% MM subset (M $\times$ M) contains tuples \texttt{(m1,m2)} such that
%method \texttt{m1} may call method \texttt{m2}.
%\item
%reachableI subset I contains the set of call sites that may be reachable.
%\item
%reachableT subset T contains the set of classes that may be reachable.
\end{itemize}
\item
{\it Points-to information:}
\begin{itemize}
\item
FH subset (F $\times$ H) contains tuples \texttt{(f,h)} such that static field
(i.e.  global variable) \texttt{f} may point to an object allocated at site
\texttt{h}.
\item
VH subset (V $\times$ H) contains tuples \texttt{(v,h)} such that local variable
\texttt{v} may point to an object allocated at site \texttt{h}.
\item
HFH subset (H $\times$ F $\times$ H) contains tuples \texttt{(h1,f,h2)} such
that instance field \texttt{f} of some object allocated at site \texttt{h1} may
point to some object allocated at site \texttt{h2}.
\end{itemize}
\end{itemize}

\subsection{Context-Sensitive Analysis}

In a context-sensitive analysis, there is no longer just one abstract memory
location per object allocation site.  Rather, the set of objects a reference can
point to depends on the \textit{context} in which the method containing the
reference is called.  Whereas a context-insensitive analysis talks about the
domain of methods (M) and the domain of allocation sites (H), a
context-sensitive analysis talks about the domain of abstract contexts, labeled
C.  Elements of domain C contain both abstract calling contexts and abstract
objects. (These are merged for reasons described below.)

Chord has several context-sensitive analyses, but they all expose the same
relations, which are described below:

\begin{itemize}
\item
{\it Context information:}
\begin{itemize}
\item
C is the domain of all abstract calling contexts and abstract objects.  Each
element in this domain is a sequence of zero or more sites, where each site can
be a call site or an object allocation site.  A sequence may have mixed call
sites and object allocation sites.  The most significant site (i.e. the first
site) in each sequence is called the head; the remaining sub-sequence is called
the tail.  The below three relations relate a sequence with its head and tail.
\item
CC subset (C $\times$ C) contains tuples \texttt{(c1,c2)} such that context
\texttt{c2} is the tail of context \texttt{c1}.
\item
CH subset (C $\times$ H) contains tuples \texttt{(c,h)} such that object
allocation site \texttt{h} is the head of context \texttt{c}.
\item
CI subset (C $\times$ I) contains tuples \texttt{(c,i)} such that call site
\texttt{i} is the head of context \texttt{c}.
\end{itemize}
\item
{\it Call-graph information:}
\begin{itemize}
\item
rootCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that method
\texttt{m} is an entry method in context \texttt{c}.
\item
reachableCM subset (C $\times$ M) contains tuples \texttt{(c,m)} such that
method \texttt{m} may be reachable in context \texttt{c}.
\item
CICM subset (C $\times$ I $\times$ C $\times$ M) contains tuples
\texttt{(c1,i,c2,m)} such that call site \texttt{i} in context \texttt{c1} may
call method \texttt{m2} in context \texttt{c2}.
\end{itemize}
\item
{\it Points-to information:}
\begin{itemize}
\item
FC subset (F $\times$ C) contains tuples \texttt{(f,o)} such that static field
(i.e. global variable) \texttt{f} may point to object \texttt{o}.
\item
CVC subset (C $\times$ V $\times$ C) contains tuples \texttt{(c,v,o)} such that
local variable \texttt{v} may point to object \texttt{o} in context \texttt{c}
of that variable's declaring method.  Note that both \texttt{o} and \texttt{c}
are elements of domain C.
\item
CFC subset (C $\times$ F $\times$ C) contains tuples \texttt{(o1,f,o2)} such
that instance field \texttt{f} of object \texttt{o1} may point to object
\texttt{o2}.  \end{itemize}
\end{itemize}

{\bf Under Construction: explain object-sensitive vs context-sensitive, plus how
to invoke.}

\section{Static Datarace Analysis}

To run Chord's static datarace analysis, run the following command:

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=datarace-java run
\end{verbatim}
\end{framed}

where \code{<WORK_DIR>} is a directory containing a file named
\code{chord.properties} that defines properties \code{chord.main.class},
\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
the application classpath, and the Java source path, respectively, of the
program to be analyzed.

Directory \code{<CHORD_MAIN_DIR>/examples/datarace_test/} provides a toy Java
program on which one can run the datarace analysis.  First run {\tt ant} in that
directory (in order to compile the program's {\tt .java} files to {\tt .class}
files) and then run the above command with \code{<WORK_DIR>} replaced by
\code{examples/datarace_test/}.  Upon successful completion, the following files
should be produced in directory \code{examples/datarace_test/chord_output/}:

\begin{itemize}
\item
File \code{dataraces_by_fld.html}, listing all dataraces grouped by the field on
which they occur; all dataraces on the same instance field or the same static
field are listed in the same group, and so are all dataraces on array elements.
\item
File \code{dataraces_by_obj.html}, listing all dataraces grouped by the abstract
object on whose field they occur; dataraces on all static fields are listed in
the same group, and so are dataraces on different instance fields of the same
abstract object.
\end{itemize}

\section{Static Deadlock Analysis}

To run Chord's static deadlock analysis, run the following command:

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> -Dchord.run.analyses=deadlock-java run
\end{verbatim}
\end{framed}

where \code{<WORK_DIR>} is a directory containing a file named
\code{chord.properties} that defines properties \code{chord.main.class},
\code{chord.class.path}, and \code{chord.src.path} specifying the main class,
the application classpath, and the Java source path, respectively, of the
program to be analyzed.

Directory \code{<CHORD_MAIN_DIR>/examples/deadlock_test/} provides a toy Java
program on which one can run the deadlock analysis.  First run {\tt ant} in that
directory (in order to compile the program's {\tt .java} files to {\tt .class}
files) and then run the above command with \code{<WORK_DIR>} replaced by
\code{examples/deadlock_test/}.  Upon successful completion, the file
\code{deadlocks.html} should be produced in directory
\code{examples/deadlock_test/chord_output/}.